### 2018-10-05 neu_torque.tex
1

%% ****** Start of file apstemplate.tex ****** %
%%
%%
%%   This file is part of the APS files in the REVTeX 4 distribution.
%%   Version 4.1r of REVTeX, August 2010
%%
%%
%%   Copyright (c) 2001, 2009, 2010 The American Physical Society.
%%
%%   See the REVTeX 4 README file for restrictions and more information.
%%
%
% This is a template for producing manuscripts for use with REVTEX 4.0
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.
%
% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
%  Add 'draft' option to mark overfull boxes with black boxes
%  Add 'showpacs' option to make PACS codes appear
%  Add 'showkeys' option to make keywords appear
%\documentclass[aps,prl,preprint,groupedaddress]{revtex4-1}
%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4-1}
%\documentclass[aps,prl,reprint,groupedaddress]{revtex4-1}
\documentclass[aps,prd,twocolumn,amsmath,amssymb,amsfont,superscriptaddress]{revtex4-1}
\usepackage{amsmath,aas_macros,xcolor,graphicx}
% You should use BibTeX and apsrev.bst for references
% Choosing a journal automatically selects the correct APS
% BibTeX style file (bst file), so only uncomment the line
% below if necessary.
%\bibliographystyle{apsrev4-1}
\bibliographystyle{h-physrev3}

\newcommand{\bs}{\boldsymbol}
\newcommand{\diff}{{\mathrm d}}
\newcommand{\Cov}{\mathsf{C}}
\newcommand{\Fish}{\mathsf{F}}
\newcommand{\Cl}{\mathcal {C}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Msun}{M_\odot}

\newcommand{\tcb}{\textcolor{blue}}
\newcommand{\tcv}{\textcolor{violet}}
\newcommand{\tcr}{\textcolor{red}}
\newcommand{\tcx}{\textcolor{teal}}

\begin{document}

% Use the \preprint command to place your local institutional report
% number in the upper righthand corner of the title page in preprint mode.
% Multiple \preprint commands are allowed.
% Use the 'preprintnumbers' class option to override journal defaults
% to display numbers if necessary
%\preprint{}

%Title of paper
\title{Neutrino Torque}

% repeat the \author .. \affiliation  etc. as needed
% \email, \thanks, \homepage, \altaffiliation all apply to the current
% author. Explanatory text should go in the []'s, actual e-mail
% address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro foreach each type of information

% \affiliation command applies to all authors since the last
% \affiliation command. The \affiliation command should follow the
% other information
% \affiliation can be followed by \email, \homepage, \thanks as well.
\author{Hao-Ran~Yu}\email{\url{haoran@cita.utoronto.ca}}
\affiliation{Tsung-Dao Lee Institute, Shanghai Jiao Tong University, Shanghai, 200240, China}
\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, M5S 3H8, Ontario, Canada}
\affiliation{Department of Astronomy, Shanghai Jiao Tong University, Shanghai, 200240, China}

\author{Ue-Li~Pen}\email{\url{pen@cita.utoronto.ca}}
\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, M5S 3H8, Ontario, Canada}
\affiliation{Tsung-Dao Lee Institute, Shanghai Jiao Tong University, Shanghai, 200240, China}
\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, Toronto, ON M5S 3H4, Canada}
\affiliation{Canadian Institute for Advanced Research, CIFAR Program in Gravitation and Cosmology, Toronto, ON M5G 1Z8, Canada}
\affiliation{Perimeter Institute for Theoretical Physics, Waterloo, ON N2L 2Y5, Canada}

\author{Xin~Wang}\email{\url{xwang@cita.utoronto.ca}}
\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, M5S 3H8, Ontario, Canada}

%Collaboration name if desired (requires use of superscriptaddress
%option in \documentclass). \noaffiliation is required (may also be
%used with the \author command).
%\collaboration can be followed by \email, \homepage, \thanks as well.
%\collaboration{}
%\noaffiliation

\date{\today}

\begin{abstract}
Neutrino mass is a long-standing physics problem. The oscillation experiments and cosmological observations provide respectively lower and upper bound on their sum of mass. Using the cosmic large scale structures to constrain the neutrino mass we encounter difficulties in disentangling other degeneracy parameters and cosmic variance. Here we show that the presence of neutrino mass provides a unique contribution to spin directions of galaxies. This neutrino torque effect is unaffected by galaxy bias and cosmic variance. Recent isobaric reconstruction techniques can reliably predict the tidal shear field of neutrinos and the resulting neutrino torque can be predicted. Upcoming galaxy surveys are promising to use the neutrino torque effect to independently constrain the neutrino mass.

\end{abstract}

% insert suggested PACS numbers in braces on next line
\pacs{}
% insert suggested keywords - APS authors don't need to do this
%\keywords{}

%\maketitle must follow title, authors, abstract, \pacs, and \keywords
\maketitle

\tcb{\textit{Introduction.---}}
The flavour oscillation experiments \citep{2002PhRvL..89a1301A} discovered the mass splittings of neutrinos and placed a lower bound of the sum of their mass $M_\nu \equiv \sum_{i=1}^3 m_{\nu_i} \gtrsim$ 0.05 eV \citep{2014ChPhC..38i0001O}. 
The existence of neutrino mass has profound impacts on cosmic evolution. 
In the early Universe relativistic neutrinos modulate the matter-to-radiation ratio, and based on which the cosmic microwave background observations provide an upper bound of $M_\nu\lesssim 0.23$ eV \citep{2016A&A...594A..13P}. 
In the late Universe, neutrinos become non-relativistic, and contribute to the matter energy density $\Omega_m$. 
Unlike the majority of matter, the cold dark matter (CDM) and baryons, neutrinos maintain a high velocity dispersion, known as ``free streaming'', which reduces their gravitational collapse on small scales. A number of large scale structure (LSS) surveys 
(such as Euclid \cite{2011arXiv1110.3193L},
LSST \cite{2009arXiv0912.0201L}, eBOSS \cite{2016AJ....151...44D},
WFIRST\cite{2012SPIE.8442E..1UG}, and DESI \cite{2015AAS...22533605E}) 
are aimed to improve this upper bound using neutrino effects on LSS,
however we usually encounter difficulties in disentangling neutrino effects from other parameters, and in understanding halo bias and cosmic variance.
Recently, new nonlinear neutrinos effects on LSS are proposed to give independent measurements on neutrino mass \citep{2014PhRvL.113m1301Z,2016PhRvL.116n1301Z,2017NatAs...1E.143Y}.
Here we present a novel neutrino torque effect which modulates the angular momenta (spin) of dark matter halos, which can be measured to constrain the neutrino mass.

\tcb{\textit{Theory.---}}
In the picture of LSS formation, gravitational instability lets initial density fluctuations form dark matter halos, in which galaxies are embedded. While in these highly nonlinear structures, uncertainties of halo bias, halo merging history and baryonic mechanisms obstruct us from clearly understanding the statistics like number counts and morphologies, the spins of galaxies/halos represent a clean structure formation history in the linear Universe, which are poorly appreciated.

The spin of a halo in Lagrangian space $\bs{j}_L\propto\int_{V_L}\bs{q}\times\bs{\nabla}\phi\diff^3\bs{q}$ can be written in first order as $\bs{j}_I\propto\bs{\epsilon}\,\bold{I}\,\bold{T}$ \citep{1984ApJ...286...38W}, where $\phi$ is the gravitational potential, $\bs{q}$ is the Lagrangian coordinates relative to the center of mass of the protohalo in volume $V_L$, $\bold{I}=(I_{ij})=(\int_{V_L} q_i q_j \diff^3\bs{q})$ is the inertia tensor, $\bold{T}=(T_{ij})=(\partial_i\partial_j\phi)$ is the tidal shear tensor, and $\bs{\epsilon}=(\epsilon_{ijk})$ is the Levi-Civita symbol. This picture shows that the initial spin of the protohalo is generated by the misalignment between the protogalactic inertia tensor and local gravitational shear tensor, and grows at first order in the early Universe \citep{1984ApJ...286...38W}. Actually, as halos collapse, nonlinear gravity field is hard to torque small objects, and we expect the spin directions are contributed mostly in the linear regime. These can be straightforwardly tested in numerical simulations. The halo merging and baryonic feedback, usually been discussed in the halo/galaxy alignment studies, do not affect the angular momentum conservation, because connected/disconnected matter and baryonic matter in Lagrangian space are torqued by the same gravitational shear.

Neutrino density field traces CDM on large scales while smoothly distributed on small scales, because the free streaming of neutrinos makes them less susceptible to the total gravitational potential. This contributes to a predictable tidal tensor $\bold{T}_\nu(m_{\nu_i})$ depending on the their mass. Acting on the protogalactic inertia tensor $\bold{I}$ it generates an additional neutrino torque $\bs{j}^I_\nu\propto\bs{\epsilon}\,\bold{I}\,\bold{T}_\nu$. The inertia tensor $\bold{I}$ is also predictable by using the fact that protogalactic inertia $\bold{I}$ and the tidal field from CDM $\bold{T}_c$ are highly correlated in Lagrangian space \citep{2000ApJ...532L...5L,2001ApJ...555..106L}, which can be understood that $\bold{I}$ is a collection of matter to be shell-crossed to form a halo, being parallel with $\bold{T}_c$.

To optimize the cross-correlation, we construct an equivalent inertia $\bold{I}_r$ from $\bold{T}_c$, in order that $\bs{\epsilon}\,\bold{I}_r\bold{T}_\nu$ is maximumly correlated with the integrated neutrino torque 
\begin{equation}\label{eq.jnu0}
\bs{j}^I_{\nu 0}=\int a(\tau)D(\tau)\bs{\epsilon}\bold{I}(\tau)\bold{T}_\nu(\tau)\diff\tau.
\end{equation}
In the primary coordinate of $\bold{T}_c$, $\bold{T}_c$ can be eigen-decomposed as $\bold{T}_c=\sum_{i=1}^3\bold{T}^{\lambda_i}_c$ and find $\alpha_i$ such that the reconstructed
\begin{equation}\label{eq.Ir}
	\bold{I}_r=\sum_{i=1}^3\alpha_i\bold{T}^{\lambda_i}_c
\end{equation}
optimizes the cross-correlation $\mu(\bs{j}^I_{\nu 0},\bs{j}^{I_r})$, where
\begin{equation}\label{eq.jnuIr}
 	\bs{j}^{I_r}_\nu=\bs{\epsilon}\,\bold{I}_r\bold{T}_\nu
\end{equation}
is the reconstructed neutrino spin direction.

\begin{figure}
\centering
  \includegraphics[width=1\linewidth]{f1}
 \caption{Illustration of the initial spin direction $\bs{j}_L$ of a protohalo region, and its tidal torque theory prediction $\bs{j}_I$ calculated by the inertia tensor $\bold{I}$ and tidal shear tensor $\bold{T}_c$. The halo at present epoch is marked with the $r_{200}$ circle with its spin direction $\bs{j}_0$, highly correlated with $\bs{j}_L$ and $\bs{j}_I$. We show that neutrino tidal torque $\bs{j}_\nu$ is generally uncorrelated with $\bs{j}_0$.}\label{fig.1}
\end{figure}

\tcb{\textit{Simulation.---}}
These correlations and coefficients are tested across a set of high-resolution $N$-body simulations \citep{2018ApJS..237...24Y}. Given any halo produced by the simulation, all the belonging particles are mapped back to Lagrangian space. The status of this definite set of particles can be traced in a resimulation, for example, their spin $\bs{j}(z)=\sum\bs{x}(z)\times\bs{v}(z)$ relative to their center of mass. In Fig.\ref{fig.1} we plot (note that all quantities are projected onto the plane of this letter) the LSS from a simulation at redshift $z=0$, and mark one selected halo with its $r_{200}$ radius (within which the mean halo density is 200 times the mean matter density of the universe) and its spin direction $\bs{j}(z=0)$ (all the spin arrows are normalized to have 15 Mpc$/h$). 
We illustrate Lagrangian mapping of this halo with its protohalo's column density (the orange clouds), with its initial spin direction $\bs{j}_L$. We can see that the Lagrangian spin $\bs{j}_L$ and Eulerian spin $\bs{j}_0$ have similar directions. \tcb{Indeed, by averaging over all the halos in the simulation, we find the expectation value of the cross-correlation coefficient $\left\langle \bs{j}(z) \cdot \bs{j}_L \right\rangle$ is smoothly decreasing from unity to 0.80.} This shows that the halo spin direction is conserved from the initial state all the way to the final state of the Universe.

We next examine the tidal torque formulation. To the first order approximation, the protohalo's inertia tensor $\bold{I}$ is represented by an equivalent ellipsoid (projected as a solid ellipse in Fig.\ref{fig.1}) with its volume equals to $V_L$.
Similarly, the initial tidal shear tensor from CDM, $\bold{T}_c$, is shown the dotted green ellipsoid with same volume. The initial tidal shear from neutrinos, $\bold{T}_\nu$ is shown by the dotted red ellipsoid (8 times the volume of $V_L$, for clearity). The initial spins given by the tidal torque theory are the interplays between $\bold{I}$ and $\bold{T}_{c/\nu}$, shown in Fig.\ref{fig.1}) as $\bs{j}_I$ and $\bs{j}_\nu$. Obviously $\bs{j}_I$ has similar direction with $\bs{j}_L$ and $\bs{j}(z=0)$, where as $\bs{j}_\nu$ gives an independent direction. \tcb{From the initial condition to the end of the simulation, $\left\langle \bs{j}(z) \cdot \bs{j}_I \right\rangle$ decreases smoothly from 0.75 to 0.69. For neutrinos, the first-order tidal torque approximation gives a perfect (with cross-correlation 0.99) representation of the actual neutrino torque and it has generally less than 0.2 cross-correlation with CDM torques.} This is expected in that $\bold{T}_c$ dominates locally \tcr{(power spectrum index?)} whereas $\bold{T}_\nu$ is contributed beyond the neutrino free-streaming scale \tcr{(is this a 4-point function?)}. These two species, however, have a highly correlated contribution in structure formation. When we consider the forces that the two species exerted to the protohalo, $\bs{F}_{c/\nu}\propto\int_{V_L}\bs{\nabla}\phi_{c/\nu}\diff^3\bs{q}$, the cross-correlation between two species is as high as 0.86 \tcr{(power spectrum index?)}.

By Eq.(\ref{eq.jnu0}) we estimate the magnitude of integrated neutrino torque \tcb{$\left\langle|\bs{j}_{0\nu}|/|\bs{j}_0|\right\rangle\simeq 3\times10^{-4}$}. In particular, the effect given by the smoother distribution for neutrinos relative to CDM accounts 0.03, while the neutrino fraction $f_\nu=3.5\times 10^{-3}$ (for $M_\nu=0.05$ eV) and the backreaction from neutrinos to CDM \tcr{$\sqrt{8}$} contribute the rest.

\tcb{Measured from simulations, $(\alpha_1,\alpha_2,\alpha_3)=(-0.18,0.18,0.02)$ optimize the cross-correlation $\left\langle\bs{j}^I_{\nu 0}\cdot\bs{j}^{I_r}_\nu\right\rangle=0.19$. In this case we need $4\times 10^8$ halos to have a $1\sigma$ detection (see Appendix A)}.

\tcb{\textit{Discussion.---}} 
We notice that the coefficients are prominently higher than \citep{2000ApJ...532L...5L}, and we investigate numerical errors that may affect the results. P3M (particle-particle particle-mesh) algorithms result in higher $\mu_L$ and $\mu_I$ compared to PM (particle-mesh), where additional tangential forces in PM violate the angular momentum conservation. Higher mass halos in a given simulation generally have slightly higher $\mu_L$ and $\mu_I$, however the correlation is enhanced greatly as we use higher mass resolutions.

\tcb{\textit{Conclusion.---}}

% body of paper here - Use proper section commands
% References should be done using the \cite, \ref, and \label commands
%\section{}
% Put \label in argument of \section for cross-referencing
%\section{\label{}}
%\subsection{}
%\subsubsection{}

% If in two-column mode, this environment will change to single-column
% format so that long equations can be displayed. Use
% sparingly.
%\begin{widetext}
% put long equation here
%\end{widetext}

% figures should be put into the text as floats.
% Use the graphics or graphicx packages (distributed with LaTeX2e)
% and the \includegraphics macro defined in those packages.
% See the LaTeX Graphics Companion by Michel Goosens, Sebastian Rahtz,
% and Frank Mittelbach for instance.
%
% Here is an example of the general form of a figure:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Use the figure* environment if the figure should span across the
% entire page. There is no need to do explicit centering.

% \begin{figure}
% \includegraphics{}%
% \caption{\label{}}
% \end{figure}

% Surround figure environment with turnpage environment for landscape
% figure
% \begin{turnpage}
% \begin{figure}
% \includegraphics{}%
% \caption{\label{}}
% \end{figure}
% \end{turnpage}

% tables should appear as floats within the text
%
% Here is an example of the general form of a table:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
% \begin{tabular}{} command.
% The ruledtabular enviroment adds doubled rules to table and sets a
% reasonable default table settings.
% Use the table* environment to get a full-width table in two-column
% Add \usepackage{longtable} and the longtable (or longtable*}
% environment for nicely formatted long tables. Or use the the [H]
% placement option to break a long table (with less control than 
% in longtable).
% \begin{table}%[H] add [H] placement to break table across pages
% \caption{\label{}}
% \begin{ruledtabular}
% \begin{tabular}{}
% Lines of table here ending with \\
% \end{tabular}
% \end{ruledtabular}
% \end{table}

% Surround table environment with turnpage environment for landscape
% table
% \begin{turnpage}
% \begin{table}
% \caption{\label{}}
% \begin{ruledtabular}
% \begin{tabular}{}
% \end{tabular}
% \end{ruledtabular}
% \end{table}
% \end{turnpage}

% Specify following sections are appendices. Use \appendix* if there
% only one appendix.
\appendix
\section{How many halos do we need in 3D}
There are $N$ halos with their unit spin vector randomly distributed on a 2D sphere, $|\bs{j}|=1$ and $\left\langle \bs{j} \right\rangle =\bs{0}$. Adding an additional vector $a\hat{\bs{x}}$ ($a\ll 1$) to $\bs{j}$ and normalize we have $\bs{j}'=(\bs{j}+a\hat{\bs{x}})/|\bs{j}+a\hat{\bs{x}}|$,then we project $\bs{j}'$ onto $\hat{\bs{x}}$ and we get the signal $b=\bs{j}'\cdot\hat{\bs{x}}$. The numerical statistics are as following: $\left\langle b \right\rangle=2a/3$ and $\sigma(b)=1/\sqrt{3N}$. So, to get a $n\sigma$ detection, $N=3n^2/4a^2$ halos will be needed.

% If you have acknowledgments, this puts in the proper section head.
%\begin{acknowledgments}
% put your acknowledgments here.
%\end{acknowledgments}

% Create the reference section using BibTeX:
\bibliography{haoran_ref}

\end{document}
%
% ****** End of file apstemplate.tex ******

### 2018-10-09 neu_torque.tex
added figure

@@
 
-The spin of a halo in Lagrangian space $\bs{j}_L\propto\int_{V_L}\bs{q}\times\bs{\nabla}\phi\diff^3\bs{q}$ can be written in first order as $\bs{j}_I\propto\bs{\epsilon}\,\bold{I}\,\bold{T}$ \citep{1984ApJ...286...38W}, where $\phi$ is the gravitational potential, $\bs{q}$ is the Lagrangian coordinates relative to the center of mass of the protohalo in volume $V_L$, $\bold{I}=(I_{ij})=(\int_{V_L} q_i q_j \diff^3\bs{q})$ is the inertia tensor, $\bold{T}=(T_{ij})=(\partial_i\partial_j\phi)$ is the tidal shear tensor, and $\bs{\epsilon}=(\epsilon_{ijk})$ is the Levi-Civita symbol.
+The spin of a halo in Lagrangian space $\bs{j}_L\propto\int_{V_L}\bs{q}\times-\bs{\nabla}\phi\diff^3\bs{q}$ can be written in first order as $\bs{j}_I\propto\bs{\epsilon}\,\bold{I}\,\bold{T}$ \citep{1984ApJ...286...38W}, where $\phi$ is the gravitational potential, $\bs{q}$ is the Lagrangian coordinates relative to the center of mass of the protohalo in volume $V_L$, $\bold{I}=(I_{ij})=(\int_{V_L} q_i q_j \diff^3\bs{q})$ is the inertia tensor, $\bold{T}=(T_{ij})=(\partial_i\partial_j\phi)$ is the tidal shear tensor, and $\bs{\epsilon}=(\epsilon_{ijk})$ is the Levi-Civita symbol.
 This picture shows that the initial spin of the protohalo is generated by the misalignment between the protogalactic inertia tensor and local gravitational shear tensor, and grows at first order in the early Universe \citep{1984ApJ...286...38W}.
@@
 Given any halo produced by the simulation, all the belonging particles are mapped back to Lagrangian space.
-The status of this definite set of particles can be traced in a resimulation, for example, their spin $\bs{j}(z)=\sum\bs{x}(z)\times\bs{v}(z)$ relative to their center of mass.
+The status of this definite set of particles can be traced in a resimulation, for example, their spin $\bs{j}(z)=m_p\sum\bs{x}(z)\times\bs{v}(z)$ relative to their center of mass.
 In Fig.\ref{fig.1} we plot (note that all quantities are projected onto the plane of this letter) the LSS from a simulation at redshift $z=0$, and mark one selected halo with its $r_{200}$ radius (within which the mean halo density is 200 times the mean matter density of the universe) and its spin direction $\bs{j}(z=0)$ (all the spin arrows are normalized to have 15 Mpc$/h$).

### 2018-10-20 neu_torque.tex
title and abstract

--- 2018-10-05 neu_torque.tex
@@
 %Title of paper
-\title{Neutrino Torque}
+\title{Parity-odd Neutrino Torque Detection}
 
@@
 \begin{abstract}
-Neutrino mass is a long-standing physics problem.
-The oscillation experiments and cosmological observations provide respectively lower and upper bound on their sum of mass.
-Using the cosmic large scale structures to constrain the neutrino mass we encounter difficulties in disentangling other degeneracy parameters and cosmic variance.
-Here we show that the presence of neutrino mass provides a unique contribution to spin directions of galaxies.
-This neutrino torque effect is unaffected by galaxy bias and cosmic variance.
-Recent isobaric reconstruction techniques can reliably predict the tidal shear field of neutrinos and the resulting neutrino torque can be predicted.
-Upcoming galaxy surveys are promising to use the neutrino torque effect to independently constrain the neutrino mass.
+Cosmological observations are promising ways for discovering the neutrino mass properties.
+The upper bound on their sum of mass is given by the cosmic microwave background and large scale structures.
+These measurements are all parity-even.
+Here we show that, the presence of neutrino mass provides a unique contribution to the directions of the angular momentum  of galaxies, which is the first parity-odd neutrino effect on large scale structures.
+This parity-odd observable is free of the contamination of the linear perturbation theory, and can be cleanly separated from other non-gravitational effects.
+A complete 21-cm neutral Hydrogen halo survey deep to redshift 1 can give a $5\sigma$ confidence level of detecting the neutrino torque effect if the sum of neutrino masses is 0.05 eV.
 
@@
 \tcb{\textit{Introduction.---}}
+Neutrino mass is a long-standing physics problem.
 The flavour oscillation experiments \citep{2002PhRvL..89a1301A} discovered the mass splittings of neutrinos and placed a lower bound of the sum of their mass $M_\nu \equiv \sum_{i=1}^3 m_{\nu_i} \gtrsim$ 0.05 eV \citep{2014ChPhC..38i0001O}.
@@
 In the picture of LSS formation, gravitational instability lets initial density fluctuations form dark matter halos, in which galaxies are embedded.
-While in these highly nonlinear structures, uncertainties of halo bias, halo merging history and baryonic mechanisms obstruct us from clearly understanding the statistics like number counts and morphologies, the spins of galaxies/halos represent a clean structure formation history in the linear Universe, which are poorly appreciated.
+While in these highly nonlinear structures, uncertainties of halo bias, halo merging history and baryonic mechanisms obstruct us from clearly understanding the statistics like number counts and morphologies \citep{2018MNRAS.tmp.2447C}, the spins of galaxies/halos represent a clean structure formation history in the linear Universe, which are poorly appreciated.
 
@@
 
-\tcb{Measured from simulations, $(\alpha_1,\alpha_2,\alpha_3)=(-0.18,0.18,0.02)$ optimize the cross-correlation $\left\langle\bs{j}^I_{\nu 0}\cdot\bs{j}^{I_r}_\nu\right\rangle=0.19$.
-In this case we need $4\times 10^8$ halos to have a $1\sigma$ detection (see Appendix A)}.
+Measured from simulations, \tcb{$(\alpha_1,\alpha_2,\alpha_3)=(-0.18,0.18,0.02)$} optimizes the cross-correlation \tcb{$\left\langle\bs{j}^I_{\nu 0}\cdot\bs{j}^{I_r}_\nu\right\rangle=0.19$}.
+In this case we need \tcb{$4\times 10^8$} halos to have a $1\sigma$ detection (see Appendix A).
+If we improve the reconstruction of $\bold{I}_R$, the lower limit of required halos is \tcb{$1.7\times 10^7$}.
 
 \tcb{\textit{Discussion.---}} 
-We notice that the coefficients are prominently higher than \citep{2000ApJ...532L...5L}, and we investigate numerical errors that may affect the results.
-P3M (particle-particle particle-mesh) algorithms result in higher $\mu_L$ and $\mu_I$ compared to PM (particle-mesh), where additional tangential forces in PM violate the angular momentum conservation.
+Accuracy and mass dependency: we notice that the cross-correlation coefficients between $\bs{j}_0$ and initial values $\bs{j}_L$ and $\bs{j}_T$ are prominently higher than what was measured in \citep{2000ApJ...532L...5L}.
+We investigate that numerical errors that may affect the results.
+P3M (particle-particle particle-mesh) algorithms result in higher cross-correlation between initial and final spins, compared to PM (particle-mesh), where additional tangential forces in PM violate the angular momentum conservation.
 Higher mass halos in a given simulation generally have slightly higher $\mu_L$ and $\mu_I$, however the correlation is enhanced greatly as we use higher mass resolutions.
+All other cross-correlation measurements have only weak dependencies on the halo mass, even in a fixed simulation.
+With different box sizes, mass resolutions, force resolutions, and find that the results are consistent across these simulations.
+Especially, in the estimation of number of halo spins needed to detect the neutrino torque, the deciding number is the cross-correlation $\left\langle\bs{j}^I_{\nu 0}\cdot\bs{j}^{I_r}_\nu\right\rangle=0.19$.
+This has a very weak dependence on halo mass and configuration of the simulation.
+
+Reconstruction: to extract the neutrino torque from observable galaxy surveys, we need to reconstruct the initial conditions.
+There are many emerging successful reconstruction algorithms \citep{2017PhRvD..96l3502Z,2017ApJ...847..110Y,2014ApJ...794...94W}, mostly based on the $E$-mode clustering in Lagrangian space \citep{2017PhRvD..95d3501Y}.
+Least action algorithms.
+
+Baryonic effects (+ etc.) in reconstruction and angular momentum conservation:
+
+Observability: surveys like the Hubble Sphere Hydrogen Survey (HSHS) \citep{2006astro.ph..6104P} can cover 1 billion galaxies.
+
+Beyond standard models: it was also suggested that neutrino mass is generated by a gravitational $\theta$-term condensation as the source of dark energy \citep{2016PhRvD..93k3002D}.
+
+Cosmic variance:
+
+Bias:
 
 \tcb{\textit{Conclusion.---}}
+
+\tcb{\textit{Acknowledgments.---}}
+We acknowledge funding from NSERC.
 
@@
 \appendix
-\section{How many halos do we need in 3D}
+\section{Errors in 3D}
 There are $N$ halos with their unit spin vector randomly distributed on a 2D sphere, $|\bs{j}|=1$ and $\left\langle \bs{j} \right\rangle =\bs{0}$.
-Adding an additional vector $a\hat{\bs{x}}$ ($a\ll 1$) to $\bs{j}$ and normalize we have $\bs{j}'=(\bs{j}+a\hat{\bs{x}})/|\bs{j}+a\hat{\bs{x}}|$,then we project $\bs{j}'$ onto $\hat{\bs{x}}$ and we get the signal $b=\bs{j}'\cdot\hat{\bs{x}}$.
+Adding an additional vector $a\hat{\bs{x}}$ ($a\ll 1$) to $\bs{j}$ and normalize we have $\bs{j}'=(\bs{j}+a\hat{\bs{x}})/|\bs{j}+a\hat{\bs{x}}|$, then we project $\bs{j}'$ onto $\hat{\bs{x}}$ and we get the signal $b=\bs{j}'\cdot\hat{\bs{x}}$.
 The numerical statistics are as following: $\left\langle b \right\rangle=2a/3$ and $\sigma(b)=1/\sqrt{3N}$.
-So, to get a $n\sigma$ detection, $N=3n^2/4a^2$ halos will be needed.
+So, to get a $n\sigma$ detection,
+\begin{equation}
+	N=3n^2/4a^2 
+\end{equation}
+halos will be needed.
 

### 2018-10-23 neu_torque.tex
abs,intro,theory,reco sections rewritten

@@
 \begin{abstract}
-Cosmological observations are promising ways for discovering the neutrino mass properties.
+Cosmological observations are promising ways to improve of our understandings of the neutrino mass properties. 
 The upper bound on their sum of mass is given by the cosmic microwave background and large scale structures. 
@@
 (such as Euclid \cite{2011arXiv1110.3193L},
-LSST \cite{2009arXiv0912.0201L}, eBOSS \cite{2016AJ....151...44D},
-WFIRST\cite{2012SPIE.8442E..1UG}, and DESI \cite{2015AAS...22533605E}) 
-are aimed to improve this upper bound using neutrino effects on LSS,
-however we usually encounter difficulties in disentangling neutrino effects from other parameters, and in understanding halo bias and cosmic variance.
+LSST \cite{2009arXiv0912.0201L}, eBOSS \cite{2016AJ....151...44D}, WFIRST\cite{2012SPIE.8442E..1UG}, and DESI \cite{2015AAS...22533605E}) are aimed to improve this upper bound using neutrino effects on LSS, however we usually encounter difficulties in disentangling neutrino effects from other parameters, and in understanding halo bias and cosmic variance.
 Recently, new nonlinear neutrinos effects on LSS are proposed to give independent measurements on neutrino mass \citep{2014PhRvL.113m1301Z,2016PhRvL.116n1301Z,2017NatAs...1E.143Y}.
-Here we present a novel neutrino torque effect which modulates the angular momenta (spin) of dark matter halos, which can be measured to constrain the neutrino mass.
+For these effects one needs to carefully separate and exclude the contaminations from non-gravitational contributions.
+The only clean, gravitational-only neutrino effect on LSS is the damping of weak gravitational lensing power spectrum \citep{2016A&A...594A..13P}.
+However the amount of information and constraining power of the lensing is limited by the projection onto the 2-dimensional sky.
+Is there a 3-dimensional measure which is cleanly gravity-driven, to measure the neutrino mass?
+
+The angular momentum. The angular momenta (spin) of dark matter halos or galaxies originated by clean gravitational interactions. 
+Unlike parity-even tidal and density fields, the linear perturbation theory (other effects and bias) cannot contaminate a parity-odd spin measurement. 
+Here we present the neutrino torque effect, to modulate the spin of dark matter halos.
 
@@
 In the picture of LSS formation, gravitational instability lets initial density fluctuations form dark matter halos, in which galaxies are embedded.
-While in these highly nonlinear structures, uncertainties of halo bias, halo merging history and baryonic mechanisms obstruct us from clearly understanding the statistics like number counts and morphologies \citep{2018MNRAS.tmp.2447C}, the spins of galaxies/halos represent a clean structure formation history in the linear Universe, which are poorly appreciated.
+In these highly nonlinear structures, uncertainties of halo bias, halo merging history and baryonic mechanisms obstruct us from clearly understanding the statistics like number counts and morphologies.
+In comparison, the spins of galaxies/halos represent a local probe of gravity, especially contributed from the linear epoch of the structure formation.
 
-The spin of a halo in Lagrangian space $\bs{j}_L\propto\int_{V_L}\bs{q}\times\bs{\nabla}\phi\diff^3\bs{q}$ can be written in first order as $\bs{j}_I\propto\bs{\epsilon}\,\bold{I}\,\bold{T}$ \citep{1984ApJ...286...38W}, where $\phi$ is the gravitational potential, $\bs{q}$ is the Lagrangian coordinates relative to the center of mass of the protohalo in volume $V_L$, $\bold{I}=(I_{ij})=(\int_{V_L} q_i q_j \diff^3\bs{q})$ is the inertia tensor, $\bold{T}=(T_{ij})=(\partial_i\partial_j\phi)$ is the tidal shear tensor, and $\bs{\epsilon}=(\epsilon_{ijk})$ is the Levi-Civita symbol.
-This picture shows that the initial spin of the protohalo is generated by the misalignment between the protogalactic inertia tensor and local gravitational shear tensor, and grows at first order in the early Universe \citep{1984ApJ...286...38W}.
-Actually, as halos collapse, nonlinear gravity field is hard to torque small objects, and we expect the spin directions are contributed mostly in the linear regime.
-These can be straightforwardly tested in numerical simulations.
-The halo merging and baryonic feedback, usually been discussed in the halo/galaxy alignment studies, do not affect the angular momentum conservation, because connected/disconnected matter and baryonic matter in Lagrangian space are torqued by the same gravitational shear.
+In the tidal torque theory, the spin of a halo in Lagrangian space $\bs{j}_L\propto\int_{V_L}\bs{q}\times\bs{\nabla}\phi\diff^3\bs{q}$ 
+can be written in first order as $\bs{j}_I\propto\bs{\epsilon}\,\bold{I}\,\bold{T}$ \citep{1984ApJ...286...38W}, 
+where $\phi$ is the gravitational potential, $\bs{q}$ is the Lagrangian coordinates relative to the center of mass of the protohalo in volume $V_L$, 
+$\bold{I}_q=(I_{ij})=(\int_{V_L} q_i q_j \diff^3\bs{q})$ is the protogalactic inertia tensor, 
+$\bold{T}=(T_{ij})=(\partial_i\partial_j\phi)$ is the local tidal shear tensor, 
+and $\bs{\epsilon}=(\epsilon_{ijk})$ is the Levi-Civita symbol to collect the asymmetric components generated by
+the misalignment between $\bold{I}_q$ and $\bold{T}$. This spin grows at first order in the early Universe \citep{1984ApJ...286...38W}. 
+Actually, as halos collapse, nonlinear gravity field is hard to torque small objects, and the spin directions are expected  to be contributed mostly in the linear regime.
+These concepts can be straightforwardly tested in numerical simulations.
+In Lagrangian space, CDM and baryonic matter are torqued by the same gravitational shear, so the spin direction is not affected by baryonic effects. 
+If a halo has a merging history, it simply collects disconnected regions in Lagrangian space but conserves their total spin.
 
-Neutrino density field traces CDM on large scales while smoothly distributed on small scales, because the free streaming of neutrinos makes them less susceptible to the total gravitational potential.
-This contributes to a predictable tidal tensor $\bold{T}_\nu(m_{\nu_i})$ depending on the their mass.
-Acting on the protogalactic inertia tensor $\bold{I}$ it generates an additional neutrino torque $\bs{j}^I_\nu\propto\bs{\epsilon}\,\bold{I}\,\bold{T}_\nu$.
-The inertia tensor $\bold{I}$ is also predictable by using the fact that protogalactic inertia $\bold{I}$ and the tidal field from CDM $\bold{T}_c$ are highly correlated in Lagrangian space \citep{2000ApJ...532L...5L,2001ApJ...555..106L}, which can be understood that $\bold{I}$ is a collection of matter to be shell-crossed to form a halo, being parallel with $\bold{T}_c$.
-
-To optimize the cross-correlation, we construct an equivalent inertia $\bold{I}_r$ from $\bold{T}_c$, in order that $\bs{\epsilon}\,\bold{I}_r\bold{T}_\nu$ is maximumly correlated with the integrated neutrino torque
+Massive neutrinos contribute sub-percent fraction of the matter ingredient, and their unique spatial distribution and evolution should contribute a unique torque to halos and galaxies.
+In particular, neutrino density field traces CDM on large scales while their small scale structures are smoothed out by their high velocity dispersion, known as the free-streaming. 
+They contribute a predictable tidal tensor $\bold{T}_\nu(m_{\nu_i})$ depending on the their mass. 
+The interplay between $\bold{I}_q$ and $\bold{T}_\nu$ leads to an additional neutrino torque $\bs{j}^I_\nu\propto\bs{\epsilon}\,\bold{I}_q\,\bold{T}_\nu$.
+Integrated neutrino torque over cosmic evolution, we get the neutrino modulation of the halo spin,
 \begin{equation}\label{eq.jnu0}
@@
 \end{equation}
+Here, $\tau$ is the Newtonian time, $D(\tau)$ is the growth factor, and other quantities are functions of time.
+
+\tcb{\textit{Reconstruction.---}} 
+The feasibility of prediction of the neutrino tidal torque relies on the precision reconstruction of the cosmic evolution history. 
+There are many emerging reconstruction methods. 
+For example, the isobaric halo reconstruction can recover the initial conditions of the universe from spatial distribution of halos. 
+In the case of high number densities the reconstructed density field is correlated with the true initial density field on scales $k\lesssim 0.7 h/{\rm Mpc}$ \citep{2017ApJ...847..110Y}, 
+close to the limit ($k\lesssim 1 h/{\rm Mpc}$) of isobaric reconstruction from using direct CDM density field \citep{2017PhRvD..96l3502Z}, 
+or the limit of reconstruction from the true displacement field \citep{2017PhRvD..95d3501Y}. 
+ELUCID simulation are able to reconstruct the local Universe \citep{2014ApJ...794...94W}.
+These reconstruction techniques enable us to study the tidal field from both CDM and neutrinos in an unprecedented precision, at different epochs of the cosmic evolution. 
+Neutrinos share the same Fourier phases of CDM, reconstructed from halo reconstruction, but only differ by the ratio of their linear transfer function, depending on the neutrino masses.
+
+To apply Eq.(\ref{eq.jnu0}), $\bold{I}_q$ is also required to be reconstructed. 
+Firstly, we note the fact that, during the initial conditions, the $\bold{I}_q$ and $\bold{T}_c$ are highly correlated in Lagrangian space \citep{2000ApJ...532L...5L,2001ApJ...555..106L}, 
+which can be understood that $\bold{I}_q$ is a collection of matter to be shell-crossed to form a halo, being parallel with $\bold{T}_c$.
+Secondly, in the linear, intermediate epochs, $\bold{I}_q$ is firstly reshaped according to the dominating $\bold{T}_c$, thus the linear evolved $\bold{T}_\nu(\tau)$ will act on an evolved $\bold{I}(\tau)$.
+Further, $\bold{I}(\tau)$ will be affected by nonlinear effects, which are more difficult to predict, however halos are relatively small and their spin directions can hardly be changed dramatically.
+\tcb{A deeply study of the second step above is beyond the scope of this letter. Here we show that, even simplified to the first point above, the total neutrino torque can still be reconstructed.}
+
+We construct an equivalent inertia $\bold{I}_r$ from $\bold{T}_c$, in order that $\bs{\epsilon}\,\bold{I}_r\bold{T}_\nu$ is maximumly correlated with $\bs{j}^I_{\nu 0}$ from Eq.(\ref{eq.jnu0}).
 In the primary coordinate of $\bold{T}_c$, $\bold{T}_c$ can be eigen-decomposed as $\bold{T}_c=\sum_{i=1}^3\bold{T}^{\lambda_i}_c$ and find $\alpha_i$ such that the reconstructed
@@
   \includegraphics[width=1\linewidth]{f1}
- \caption{Illustration of the initial spin direction $\bs{j}_L$ of a protohalo region, and its tidal torque theory prediction $\bs{j}_I$ calculated by the inertia tensor $\bold{I}$ and tidal shear tensor $\bold{T}_c$.
-The halo at present epoch is marked with the $r_{200}$ circle with its spin direction $\bs{j}_0$, highly correlated with $\bs{j}_L$ and $\bs{j}_I$.
+ \caption{Illustration of the initial spin direction $\bs{j}_L$ of a typical protohalo region, and its tidal torque theory prediction $\bs{j}_I$ calculated by the inertia tensor $\bold{I}_q$ and tidal shear tensor $\bold{T}_c$.
+The halo at present epoch is marked with the $\bold{I}_{200}$ inertia tensor with its spin direction $\bs{j}_0$, highly correlated with $\bs{j}_L$ and $\bs{j}_I$.
 We show that neutrino tidal torque $\bs{j}_\nu$ is generally uncorrelated with $\bs{j}_0$.}\label{fig.1}
@@
 The status of this definite set of particles can be traced in a resimulation, for example, their spin $\bs{j}(z)=\sum\bs{x}(z)\times\bs{v}(z)$ relative to their center of mass. 
-In Fig.\ref{fig.1} we plot (note that all quantities are projected onto the plane of this letter) the LSS from a simulation at redshift $z=0$, and mark one selected halo with its $r_{200}$ radius (within which the mean halo density is 200 times the mean matter density of the universe) and its spin direction $\bs{j}(z=0)$ (all the spin arrows are normalized to have 15 Mpc$/h$).
-We illustrate Lagrangian mapping of this halo with its protohalo's column density (the orange clouds), with its initial spin direction $\bs{j}_L$.
+In Fig.\ref{fig.1} (note that all quantities are projected onto the plane of this letter), 
+we select a very massive halo ($7.8\times 10^{14}M_\odot$) to maximize the clarity of the visualization of the halo properties.
+Note that these properties (cross-correlations) have only weak dependence on halo mass. 
+The background LSS at redshift $z=0$ has the thickness $2r_{200}$ with $r_{200}$ being the halo radius within which the mean halo density is 200 times the mean matter density of the universe.
+The arrow $\bs{j}_0$ shows its spin (all the spin arrows are normalized to have 15 Mpc$/h$).
+We illustrate Lagrangian mapping of this halo with its protohalo's column density (the orange clouds), with its initial spin $\bs{j}_L$. 
+The two solid ellipsoids $\bold{I}_q$ and $\bold{I}_q$
+
  We can see that the Lagrangian spin $\bs{j}_L$ and Eulerian spin $\bs{j}_0$ have similar directions.
@@
 The initial spins given by the tidal torque theory are the interplays between $\bold{I}$ and $\bold{T}_{c/\nu}$, shown in Fig.\ref{fig.1}) as $\bs{j}_I$ and $\bs{j}_\nu$. 
-Obviously $\bs{j}_I$ has similar direction with $\bs{j}_L$ and $\bs{j}(z=0)$, where as $\bs{j}_\nu$ gives an independent direction.
+Typically $\bs{j}_I$ has similar direction with $\bs{j}_L$ and $\bs{j}(z=0)$, where as $\bs{j}_\nu$ gives an independent direction.
 \tcb{From the initial condition to the end of the simulation, $\left\langle \bs{j}(z) \cdot \bs{j}_I \right\rangle$ decreases smoothly from 0.75 to 0.69.
@@
 
-Observability: surveys like the Hubble Sphere Hydrogen Survey (HSHS) \citep{2006astro.ph..6104P} can cover 1 billion galaxies.
+Observability: surveys like the Hubble Sphere Hydrogen Survey (HSHS) \citep{2006astro.ph..6104P} can cover 1 billion galaxies. There are xxx HI galaxies with redshift 1 \citep{2004MNRAS.350.1210Z}.
 
@@
 Bias:
+
+cite later \citep{2018MNRAS.tmp.2447C}.
 

### 2018-10-24 neu_torque.tex
simulation section

@@
 \newcommand{\Fish}{\mathsf{F}}
-\newcommand{\Cl}{\mathcal {C}}
+\newcommand{\Cl}{\mathcal{C}}
 \newcommand{\R}{\mathcal{R}}
-\newcommand{\T}{\mathcal{T}}
+\newcommand{\T}{\bold{T}}
+\newcommand{\I}{\bold{I}}
+\newcommand{\spin}{\bs{j}}
 \newcommand{\Msun}{M_\odot}
@@
 Neutrino mass is a long-standing physics problem. 
-The flavour oscillation experiments \citep{2002PhRvL..89a1301A} discovered the mass splittings of neutrinos and placed a lower bound of the sum of their mass $M_\nu \equiv \sum_{i=1}^3 m_{\nu_i} \gtrsim$ 0.05 eV \citep{2014ChPhC..38i0001O}.
+The flavour oscillation experiments \citep{2002PhRvL..89a1301A} discovered the mass 
+splittings of neutrinos and placed a lower bound of the sum of their mass 
+$M_\nu \equiv \sum_{i=1}^3 m_{\nu_i} \gtrsim$ 0.05 eV \citep{2014ChPhC..38i0001O}. 
 The existence of neutrino mass has profound impacts on cosmic evolution. 
-In the early Universe relativistic neutrinos modulate the matter-to-radiation ratio, and based on which the cosmic microwave background observations provide an upper bound of $M_\nu\lesssim 0.23$ eV \citep{2016A&A...594A..13P}.
-In the late Universe, neutrinos become non-relativistic, and contribute to the matter energy density $\Omega_m$. 
-Unlike the majority of matter, the cold dark matter (CDM) and baryons, neutrinos maintain a high velocity dispersion, known as ``free streaming'', which reduces their gravitational collapse on small scales.
+In the early Universe relativistic neutrinos modulate the matter-to-radiation ratio, 
+and based on which the cosmic microwave background observations provide an 
+upper bound of $M_\nu\lesssim 0.23$ eV \citep{2016A&A...594A..13P}. 
+In the late Universe, neutrinos become non-relativistic, 
+and contribute to the matter energy density $\Omega_m$. 
+Unlike the majority of matter, the cold dark matter (CDM) and baryons, 
+neutrinos maintain a high velocity dispersion, 
+known as ``free streaming'', which reduces their gravitational collapse on small scales. 
 A number of large scale structure (LSS) surveys 
 (such as Euclid \cite{2011arXiv1110.3193L},
-LSST \cite{2009arXiv0912.0201L}, eBOSS \cite{2016AJ....151...44D}, WFIRST\cite{2012SPIE.8442E..1UG}, and DESI \cite{2015AAS...22533605E}) are aimed to improve this upper bound using neutrino effects on LSS, however we usually encounter difficulties in disentangling neutrino effects from other parameters, and in understanding halo bias and cosmic variance.
+LSST \cite{2009arXiv0912.0201L}, eBOSS \cite{2016AJ....151...44D}, WFIRST\cite{2012SPIE.8442E..1UG}, 
+and DESI \cite{2015AAS...22533605E}) are aimed to improve this upper bound using neutrino effects on LSS, 
+however we usually encounter difficulties in disentangling neutrino effects from other parameters, 
+and in understanding halo bias and cosmic variance. 
 Recently, new nonlinear neutrinos effects on LSS are proposed to give independent measurements on neutrino mass \citep{2014PhRvL.113m1301Z,2016PhRvL.116n1301Z,2017NatAs...1E.143Y}. 
@@
 In the picture of LSS formation, gravitational instability lets initial density fluctuations form dark matter halos, in which galaxies are embedded.
-In these highly nonlinear structures, uncertainties of halo bias, halo merging history and baryonic mechanisms obstruct us from clearly understanding the statistics like number counts and morphologies.
-In comparison, the spins of galaxies/halos represent a local probe of gravity, especially contributed from the linear epoch of the structure formation.
+In these highly nonlinear structures, uncertainties of halo bias, 
+halo merging history and baryonic mechanisms obstruct us from clearly 
+understanding the statistics like number counts and morphologies.
+In comparison, the spins of galaxies/halos represent a local probe of gravity, 
+especially contributed from the linear epoch of the structure formation.
 
-In the tidal torque theory, the spin of a halo in Lagrangian space $\bs{j}_L\propto\int_{V_L}\bs{q}\times\bs{\nabla}\phi\diff^3\bs{q}$ 
-can be written in first order as $\bs{j}_I\propto\bs{\epsilon}\,\bold{I}\,\bold{T}$ \citep{1984ApJ...286...38W}, 
+In the tidal torque theory, the spin of a halo in Lagrangian space $\spin_L\propto\int_{V_L}\bs{q}\times\bs{\nabla}\phi\diff^3\bs{q}$ 
+can be written in first order as $\spin_I\propto\bs{\epsilon}\,\I\,\T$ \citep{1984ApJ...286...38W}, 
 where $\phi$ is the gravitational potential, $\bs{q}$ is the Lagrangian coordinates relative to the center of mass of the protohalo in volume $V_L$, 
-$\bold{I}_q=(I_{ij})=(\int_{V_L} q_i q_j \diff^3\bs{q})$ is the protogalactic inertia tensor, 
-$\bold{T}=(T_{ij})=(\partial_i\partial_j\phi)$ is the local tidal shear tensor, 
+$\I_q=(I_{ij})=(\int_{V_L} q_i q_j \diff^3\bs{q})$ is the protogalactic inertia tensor, 
+$\T=(T_{ij})=(\partial_i\partial_j\phi)$ is the local tidal shear tensor, 
 and $\bs{\epsilon}=(\epsilon_{ijk})$ is the Levi-Civita symbol to collect the asymmetric components generated by
-the misalignment between $\bold{I}_q$ and $\bold{T}$. This spin grows at first order in the early Universe \citep{1984ApJ...286...38W}. 
+the misalignment between $\I_q$ and $\T$. This spin grows at first order in the early Universe \citep{1984ApJ...286...38W}. 
 Actually, as halos collapse, nonlinear gravity field is hard to torque small objects, and the spin directions are expected  to be contributed mostly in the linear regime.
@@
 In particular, neutrino density field traces CDM on large scales while their small scale structures are smoothed out by their high velocity dispersion, known as the free-streaming. 
-They contribute a predictable tidal tensor $\bold{T}_\nu(m_{\nu_i})$ depending on the their mass. 
-The interplay between $\bold{I}_q$ and $\bold{T}_\nu$ leads to an additional neutrino torque $\bs{j}^I_\nu\propto\bs{\epsilon}\,\bold{I}_q\,\bold{T}_\nu$.
+They contribute a predictable tidal tensor $\T_\nu(m_{\nu_i})$ depending on the their mass. 
+The interplay between $\I_q$ and $\T_\nu$ leads to an additional neutrino torque $\spin^I_\nu\propto\bs{\epsilon}\,\I_q\,\T_\nu$.
 Integrated neutrino torque over cosmic evolution, we get the neutrino modulation of the halo spin,
 \begin{equation}\label{eq.jnu0}
-\bs{j}^I_{\nu 0}=\int a(\tau)D(\tau)\bs{\epsilon}\bold{I}(\tau)\bold{T}_\nu(\tau)\diff\tau.
+\spin^I_{\nu 0}=\int a(\tau)D(\tau)\bs{\epsilon}\I(\tau)\T_\nu(\tau)\diff\tau.
 \end{equation}
@@
 Neutrinos share the same Fourier phases of CDM, reconstructed from halo reconstruction, but only differ by the ratio of their linear transfer function, depending on the neutrino masses.
+The final neutrino torque can be understood as the interaction between large (neutrino free streaming) and small (collapsing) scales.
 
-To apply Eq.(\ref{eq.jnu0}), $\bold{I}_q$ is also required to be reconstructed. 
-Firstly, we note the fact that, during the initial conditions, the $\bold{I}_q$ and $\bold{T}_c$ are highly correlated in Lagrangian space \citep{2000ApJ...532L...5L,2001ApJ...555..106L}, 
-which can be understood that $\bold{I}_q$ is a collection of matter to be shell-crossed to form a halo, being parallel with $\bold{T}_c$.
-Secondly, in the linear, intermediate epochs, $\bold{I}_q$ is firstly reshaped according to the dominating $\bold{T}_c$, thus the linear evolved $\bold{T}_\nu(\tau)$ will act on an evolved $\bold{I}(\tau)$.
-Further, $\bold{I}(\tau)$ will be affected by nonlinear effects, which are more difficult to predict, however halos are relatively small and their spin directions can hardly be changed dramatically.
+To apply Eq.(\ref{eq.jnu0}), $\I_q$ is also required to be reconstructed. 
+Firstly, we note the fact that, during the initial conditions, the $\I_q$ and $\T_c$ are highly correlated in Lagrangian space \citep{2000ApJ...532L...5L,2001ApJ...555..106L}, 
+which can be understood that $\I_q$ is a collection of matter to be shell-crossed to form a halo, being parallel with $\T_c$.
+Secondly, in the linear, intermediate epochs, $\I_q$ is firstly reshaped according to the dominating $\T_c$, thus the linear evolved $\T_\nu(\tau)$ will act on an evolved $\I(\tau)$.
+Further, $\I(\tau)$ will be affected by nonlinear effects, which are more difficult to predict, however halos are relatively small and their spin directions can hardly be changed dramatically.
 \tcb{A deeply study of the second step above is beyond the scope of this letter. Here we show that, even simplified to the first point above, the total neutrino torque can still be reconstructed.}
 
-We construct an equivalent inertia $\bold{I}_r$ from $\bold{T}_c$, in order that $\bs{\epsilon}\,\bold{I}_r\bold{T}_\nu$ is maximumly correlated with $\bs{j}^I_{\nu 0}$ from Eq.(\ref{eq.jnu0}).
-In the primary coordinate of $\bold{T}_c$, $\bold{T}_c$ can be eigen-decomposed as $\bold{T}_c=\sum_{i=1}^3\bold{T}^{\lambda_i}_c$ and find $\alpha_i$ such that the reconstructed
+We construct an equivalent inertia $\I_r$ from $\T_c$, in order that $\bs{\epsilon}\,\I_r\T_\nu$ is maximumly correlated with $\spin^I_{\nu 0}$ from Eq.(\ref{eq.jnu0}).
+In the primary coordinate of $\T_c$, $\T_c$ can be eigen-decomposed as $\T_c=\sum_{i=1}^3\T^{\lambda_i}_c$ and find $\alpha_i$ such that the reconstructed
 \begin{equation}\label{eq.Ir}
-	\bold{I}_r=\sum_{i=1}^3\alpha_i\bold{T}^{\lambda_i}_c
+	\I_r=\sum_{i=1}^3\alpha_i\T^{\lambda_i}_c
 \end{equation}
-optimizes the cross-correlation $\mu(\bs{j}^I_{\nu 0},\bs{j}^{I_r})$, where
+optimizes the cross-correlation $\mu(\spin^I_{\nu 0},\spin^{I_r})$, where
 \begin{equation}\label{eq.jnuIr}
- 	\bs{j}^{I_r}_\nu=\bs{\epsilon}\,\bold{I}_r\bold{T}_\nu
+ 	\spin^{I_r}_\nu=\bs{\epsilon}\,\I_r\T_\nu
 \end{equation}
@@
   \includegraphics[width=1\linewidth]{f1}
- \caption{Illustration of the initial spin direction $\bs{j}_L$ of a typical protohalo region, and its tidal torque theory prediction $\bs{j}_I$ calculated by the inertia tensor $\bold{I}_q$ and tidal shear tensor $\bold{T}_c$.
-The halo at present epoch is marked with the $\bold{I}_{200}$ inertia tensor with its spin direction $\bs{j}_0$, highly correlated with $\bs{j}_L$ and $\bs{j}_I$.
-We show that neutrino tidal torque $\bs{j}_\nu$ is generally uncorrelated with $\bs{j}_0$.}\label{fig.1}
+ \caption{Visualization of the neutrino torque. 
+ We show the LSS slice centered at a selected halo with depth twice the halo radius $r_{200}$. 
+ Equivalent ellipsoids by solid lines show the moment of inertia in Lagrangian and Eulerian space $\I_q$ and $\I_{200}$, 
+ while dotted ellipsoids show the tidal shear from CDM and neutrinos $\T_c$ and $\T_\nu$. $\spin_L$, 
+ $\spin_0$ are the Lagrangian and Eulerian halo spins, whereas $\spin_I$ is the tidal torque prediction. 
+ The initial neutrino torque is shown by $\spin_\nu$.}\label{fig.1}
 \end{figure}
-
 \tcb{\textit{Simulation.---}}
 These correlations and coefficients are tested across a set of high-resolution $N$-body simulations \citep{2018ApJS..237...24Y}. 
-Given any halo produced by the simulation, all the belonging particles are mapped back to Lagrangian space. 
-The status of this definite set of particles can be traced in a resimulation, for example, their spin $\bs{j}(z)=\sum\bs{x}(z)\times\bs{v}(z)$ relative to their center of mass. 
-In Fig.\ref{fig.1} (note that all quantities are projected onto the plane of this letter), 
+Given any halo formed in the simulation, all the belonging particles are mapped back to Lagrangian space. 
+The status of this definite set of particles can be traced in a resimulation of the exact same initial conditions.
+
+In Fig.\ref{fig.1} (all quantities are projected onto the plane of this letter), 
 we select a very massive halo ($7.8\times 10^{14}M_\odot$) to maximize the clarity of the visualization of the halo properties.
 Note that these properties (cross-correlations) have only weak dependence on halo mass. 
-The background LSS at redshift $z=0$ has the thickness $2r_{200}$ with $r_{200}$ being the halo radius within which the mean halo density is 200 times the mean matter density of the universe.
-The arrow $\bs{j}_0$ shows its spin (all the spin arrows are normalized to have 15 Mpc$/h$).
-We illustrate Lagrangian mapping of this halo with its protohalo's column density (the orange clouds), with its initial spin $\bs{j}_L$. 
-The two solid ellipsoids $\bold{I}_q$ and $\bold{I}_q$
+The background LSS at redshift $z=0$ has the thickness $2r_{200}$ with $r_{200}$ being the halo radius within which the mean halo density is 200 times the mean matter density of the Universe.
+The Lagrangian mapping of this halo is shown by the protohalo's column density with the orange clouds.
+To visualize the tidal torque theory, we plot ellipsoids equivalent to $\I_q$ and 
+$\I_{200}$, where $\I_{200}$ is the moment of inertia within $r_{200}$. 
+The ellipsoids with dotted lines correspond to $\T_c$ and $\T_\nu$, normalized such that their volumes are $V_L$ and $8V_L$ respectively. 
+As expected, $\I_q$ and $\T_c$ are aligned with their primary axes in parallel with the collapsing direction, or perpendicular to the filament below. 
+Their minor misalignment yields the tidal torque $\spin_I$ (all the spin arrows are normalized to have 15 Mpc$/h$), which is the first order approximation of the true initial spin $\spin_L$. 
+They are, in general, highly correlated with the spin of the final halo $\spin_0$.
+In comparison, the neutrino tidal shear $\T_\nu$ torques $\I_q$ in an other less correlated direction $\spin_\nu$.
 
- We can see that the Lagrangian spin $\bs{j}_L$ and Eulerian spin $\bs{j}_0$ have similar directions.
-\tcb{Indeed, by averaging over all the halos in the simulation, we find the expectation value of the cross-correlation coefficient $\left\langle \bs{j}(z) \cdot \bs{j}_L \right\rangle$ is smoothly decreasing from unity to 0.80.} This shows that the halo spin direction is conserved from the initial state all the way to the final state of the Universe.
+The validity of the tidal torque formulation is tested by an ensemble average over all halos, 
+across 3 orders of magnitude in mass range, and over simulations with different resolutions.
+the cross-correlation coefficients $\left\langle \spin(z) \cdot \spin_L \right\rangle$ and
+$\left\langle \spin(z) \cdot \spin_I \right\rangle$ smoothly decrease from 1 to 0.80, and
+from 0.75 to 0.69, respectively.
 
-We next examine the tidal torque formulation.
-To the first order approximation, the protohalo's inertia tensor $\bold{I}$ is represented by an equivalent ellipsoid (projected as a solid ellipse in Fig.\ref{fig.1}) with its volume equals to $V_L$.
-Similarly, the initial tidal shear tensor from CDM, $\bold{T}_c$, is shown the dotted green ellipsoid with same volume. 
-The initial tidal shear from neutrinos, $\bold{T}_\nu$ is shown by the dotted red ellipsoid (8 times the volume of $V_L$, for clearity). 
-The initial spins given by the tidal torque theory are the interplays between $\bold{I}$ and $\bold{T}_{c/\nu}$, shown in Fig.\ref{fig.1}) as $\bs{j}_I$ and $\bs{j}_\nu$. 
-Typically $\bs{j}_I$ has similar direction with $\bs{j}_L$ and $\bs{j}(z=0)$, where as $\bs{j}_\nu$ gives an independent direction.
-\tcb{From the initial condition to the end of the simulation, $\left\langle \bs{j}(z) \cdot \bs{j}_I \right\rangle$ decreases smoothly from 0.75 to 0.69.
-For neutrinos, the first-order tidal torque approximation gives a perfect (with cross-correlation 0.99) representation of the actual neutrino torque and it has generally less than 0.2 cross-correlation with CDM torques.} This is expected in that $\bold{T}_c$ dominates locally \tcr{(power spectrum index?)} whereas $\bold{T}_\nu$ is contributed beyond the neutrino free-streaming scale \tcr{(is this a 4-point function?)}.
+For neutrinos, the first-order tidal torque approximation gives a perfect (with cross-correlation 0.99) representation of the actual neutrino torque and it has generally less than 0.2 cross-correlation with CDM torques.
+This is expected in that $\T_c$ dominates locally \tcr{(power spectrum index?)} whereas $\T_\nu$ is contributed beyond the neutrino free-streaming scale \tcr{(is this a 4-point function?)}.
 These two species, however, have a highly correlated contribution in structure formation.
@@
 
-By Eq.(\ref{eq.jnu0}) we estimate the magnitude of integrated neutrino torque \tcb{$\left\langle|\bs{j}_{0\nu}|/|\bs{j}_0|\right\rangle\simeq 3\times10^{-4}$}.
+By Eq.(\ref{eq.jnu0}) we estimate the magnitude of integrated neutrino torque $\left\langle|\spin_{0\nu}|/|\spin_0|\right\rangle\simeq 3\times10^{-4}$.
 In particular, the effect given by the smoother distribution for neutrinos relative to CDM accounts 0.03, while the neutrino fraction $f_\nu=3.5\times 10^{-3}$ (for $M_\nu=0.05$ eV) and the backreaction from neutrinos to CDM \tcr{$\sqrt{8}$} contribute the rest.
 
-Measured from simulations, \tcb{$(\alpha_1,\alpha_2,\alpha_3)=(-0.18,0.18,0.02)$} optimizes the cross-correlation \tcb{$\left\langle\bs{j}^I_{\nu 0}\cdot\bs{j}^{I_r}_\nu\right\rangle=0.19$}.
-In this case we need \tcb{$4\times 10^8$} halos to have a $1\sigma$ detection (see Appendix A).
-If we improve the reconstruction of $\bold{I}_R$, the lower limit of required halos is \tcb{$1.7\times 10^7$}.
+Measured from simulations, $(\alpha_1,\alpha_2,\alpha_3)=(-0.18,0.18,0.02)$ in Eq.(\ref{eq.Ir}) optimizes the cross-correlation $\left\langle\spin^I_{\nu 0}\cdot\spin^{I_r}_\nu\right\rangle=0.19$.
+In this case we need $4\times 10^8$ halos to have a $1\sigma$ detection (see Appendix A).
+If we improve the reconstruction of $\I_R$, the lower limit of required halos is $1.7\times 10^7$.
 
 \tcb{\textit{Discussion.---}} 
-Accuracy and mass dependency: we notice that the cross-correlation coefficients between $\bs{j}_0$ and initial values $\bs{j}_L$ and $\bs{j}_T$ are prominently higher than what was measured in \citep{2000ApJ...532L...5L}.
+Accuracy and mass dependency: we notice that the cross-correlation coefficients between $\spin_0$ and initial values $\spin_L$ and $\spin_T$ are prominently higher than what was measured in \citep{2000ApJ...532L...5L}.
 We investigate that numerical errors that may affect the results.
@@
 With different box sizes, mass resolutions, force resolutions, and find that the results are consistent across these simulations.
-Especially, in the estimation of number of halo spins needed to detect the neutrino torque, the deciding number is the cross-correlation $\left\langle\bs{j}^I_{\nu 0}\cdot\bs{j}^{I_r}_\nu\right\rangle=0.19$.
+Especially, in the estimation of number of halo spins needed to detect the neutrino torque, the deciding number is the cross-correlation $\left\langle\spin^I_{\nu 0}\cdot\spin^{I_r}_\nu\right\rangle=0.19$.
 This has a very weak dependence on halo mass and configuration of the simulation.
-
-Reconstruction: to extract the neutrino torque from observable galaxy surveys, we need to reconstruct the initial conditions.
-There are many emerging successful reconstruction algorithms \citep{2017PhRvD..96l3502Z,2017ApJ...847..110Y,2014ApJ...794...94W}, mostly based on the $E$-mode clustering in Lagrangian space \citep{2017PhRvD..95d3501Y}.
-Least action algorithms.
 
@@
 \section{Errors in 3D}
-There are $N$ halos with their unit spin vector randomly distributed on a 2D sphere, $|\bs{j}|=1$ and $\left\langle \bs{j} \right\rangle =\bs{0}$.
-Adding an additional vector $a\hat{\bs{x}}$ ($a\ll 1$) to $\bs{j}$ and normalize we have $\bs{j}'=(\bs{j}+a\hat{\bs{x}})/|\bs{j}+a\hat{\bs{x}}|$, then we project $\bs{j}'$ onto $\hat{\bs{x}}$ and we get the signal $b=\bs{j}'\cdot\hat{\bs{x}}$.
+There are $N$ halos with their unit spin vector randomly distributed on a 2D sphere, $|\spin|=1$ and $\left\langle \spin \right\rangle =\bs{0}$.
+Adding an additional vector $a\hat{\bs{x}}$ ($a\ll 1$) to $\spin$ and normalize we have $\spin'=(\spin+a\hat{\bs{x}})/|\spin+a\hat{\bs{x}}|$, then we project $\spin'$ onto $\hat{\bs{x}}$ and we get the signal $b=\spin'\cdot\hat{\bs{x}}$.
 The numerical statistics are as following: $\left\langle b \right\rangle=2a/3$ and $\sigma(b)=1/\sqrt{3N}$.

### 2018-10-24 neu_torque.tex
5sigma

@@
 Measured from simulations, $(\alpha_1,\alpha_2,\alpha_3)=(-0.18,0.18,0.02)$ in Eq.(\ref{eq.Ir}) optimizes the cross-correlation $\left\langle\spin^I_{\nu 0}\cdot\spin^{I_r}_\nu\right\rangle=0.19$.
-In this case we need $4\times 10^8$ halos to have a $1\sigma$ detection (see Appendix A).
-If we improve the reconstruction of $\I_R$, the lower limit of required halos is $1.7\times 10^7$.
+In this case we need $5\times 10^9$ halos to have a $5\sigma$ detection (see Appendix A).
+If we improve the reconstruction of $\I_R$, the lower limit of required halos is $2\times 10^8$.
 

### 2018-10-24 neu_torque.tex
accuracy and nonlinear

@@
 This has a very weak dependence on halo mass and configuration of the simulation.
-
-Baryonic effects (+ etc.) in reconstruction and angular momentum conservation:
+An accurate study of the neutrino torque requires mass and force resolutions to cover the wide range of halo mass, a large box size ($>600\,{\rm Mpc}/h$) to account the neutrino tides at distance, and neutrino particles or fluids to calculate more precise nonlinear neutrino effects.
+These require a simulation comparable to the computing resources of TianNu \citep{2017RAA....17...85E}.
 
@@
 
-Cosmic variance:
-
-Bias:
-
-cite later \citep{2018MNRAS.tmp.2447C}.
+Codis \citep{2018MNRAS.tmp.2447C}.
 

### 2018-10-25 neu_torque.tex
conclusion

@@
 Measured from simulations, $(\alpha_1,\alpha_2,\alpha_3)=(-0.18,0.18,0.02)$ in Eq.(\ref{eq.Ir}) optimizes the cross-correlation $\left\langle\spin^I_{\nu 0}\cdot\spin^{I_r}_\nu\right\rangle=0.19$.
-In this case we need $5\times 10^9$ halos to have a $5\sigma$ detection (see Appendix A).
+We then need $5\times 10^9$ halos to have a $5\sigma$ detection (see Appendix \ref{app.error}).
 If we improve the reconstruction of $\I_R$, the lower limit of required halos is $2\times 10^8$.
@@
 \tcb{\textit{Conclusion.---}} 
+Enormous efforts has been contributed to the neutrino mass properties.
+Radio astronomic and cosmological surveys are promising low-cost experiments implemented in the Universe to probe basic physic mysteries. 
+The angular momentum is a 3-dimensional, gravity-driven, parity-odd measure, 
+and is easily observable, well modeled by the tidal torque theory, and well conserved over the cosmic evolution.
+It contains comparable amount of information as density field but poorly appreciated in application.
+The recently developed reconstruction techniques enable us to visit an-order-of-magnitude more precise initial conditions of the Universe in Lagrangian space.
 
@@
 \appendix
-\section{Errors in 3D}
+\section{Errors in 3D}\label{app.error}
 There are $N$ halos with their unit spin vector randomly distributed on a 2D sphere, $|\spin|=1$ and $\left\langle \spin \right\rangle =\bs{0}$.
@@
 % ****** End of file apstemplate.tex ******
+
+\begin{matrix}{ccc}
+  & & \\
+  & & \\
+  & &
+\end{matrix}

### 2018-10-25 neu_torque.tex
revised discussion and conclusion

@@
 \end{equation}
-optimizes the cross-correlation $\mu(\spin^I_{\nu 0},\spin^{I_r})$, where
+optimizes the cross-correlation between $\spin^I_{\nu 0}$ and $\spin^{I_r}_\nu$, where
 \begin{equation}\label{eq.jnuIr}
@@
 \tcb{\textit{Discussion.---}} 
-Accuracy and mass dependency: we notice that the cross-correlation coefficients between $\spin_0$ and initial values $\spin_L$ and $\spin_T$ are prominently higher than what was measured in \citep{2000ApJ...532L...5L}.
+We notice that the cross-correlation coefficients between $\spin_0$ and initial values $\spin_L$ and $\spin_T$ are prominently higher than what was measured in \citep{2000ApJ...532L...5L}.
 We investigate that numerical errors that may affect the results.
 P3M (particle-particle particle-mesh) algorithms result in higher cross-correlation between initial and final spins, compared to PM (particle-mesh), where additional tangential forces in PM violate the angular momentum conservation.
-Higher mass halos in a given simulation generally have slightly higher $\mu_L$ and $\mu_I$, however the correlation is enhanced greatly as we use higher mass resolutions.
+Higher mass halos in a given simulation generally have slightly higher cross-correlations between initial and final spins, however the correlation is enhanced greatly as we use higher mass resolutions.
 All other cross-correlation measurements have only weak dependencies on the halo mass, even in a fixed simulation.
@@
 
-Observability: surveys like the Hubble Sphere Hydrogen Survey (HSHS) \citep{2006astro.ph..6104P} can cover 1 billion galaxies. There are xxx HI galaxies with redshift 1 \citep{2004MNRAS.350.1210Z}.
+For surveys like the Hubble Sphere Hydrogen Survey (HSHS) \citep{2006astro.ph..6104P}, we are able to observe \tcr{$10^9$} HI galaxies ($>10^{12}M_\odot$), 
+in a cosmic volume $(4{\rm Gpc}/h)^3$ below redshift $z\simeq 1$   \citep{2004MNRAS.350.1210Z}. 
+Under the standard model of the Universe with $M_\nu=0.05$ eV, a $5\sigma$ detection will be reachable.
+Beyond standard models, e.g., the neutrino mass could be generated by a gravitational $\theta$-term condensation as the source of dark energy \citep{2016PhRvD..93k3002D}, the tidal torque history will be different and the numbers of galaxies are different to differentiate between these models.
 
-Beyond standard models: it was also suggested that neutrino mass is generated by a gravitational $\theta$-term condensation as the source of dark energy \citep{2016PhRvD..93k3002D}.
-
-Codis \citep{2018MNRAS.tmp.2447C}.
+%Codis \citep{2018MNRAS.tmp.2447C}.
 
@@
 It contains comparable amount of information as density field but poorly appreciated in application.
-The recently developed reconstruction techniques enable us to visit an-order-of-magnitude more precise initial conditions of the Universe in Lagrangian space.
+The recently developed reconstruction techniques enable us to visit an-order-of-magnitude more precise initial conditions of the Universe in Lagrangian space, 
+and thus to accurately reconstruct the neutrino torque effects presented in the letter.
 

### 2018-10-28 neu_torque.tex
include Ue-Li's change

--- 2018-10-25 neu_torque.tex
@@
 %\documentclass[aps,prl,reprint,groupedaddress]{revtex4-1}
-\documentclass[aps,prd,twocolumn,amsmath,amssymb,amsfont,superscriptaddress]{revtex4-1}
+\documentclass[aps,prd,twocolumn,amsmath,amssymb,amsfont,superscriptaddress,nofootinbib]{revtex4-1}
 \usepackage{amsmath,aas_macros,xcolor,graphicx}
@@
 \affiliation{Tsung-Dao Lee Institute, Shanghai Jiao Tong University, Shanghai, 200240, China}
-\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, M5S 3H8, Ontario, Canada}
+\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, M5S 3H8, ON, Canada}
 \affiliation{Department of Astronomy, Shanghai Jiao Tong University, Shanghai, 200240, China}
@@
 \author{Ue-Li~Pen}\email{\url{pen@cita.utoronto.ca}}
-\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, M5S 3H8, Ontario, Canada}
+\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, M5S 3H8, ON, Canada}
 \affiliation{Tsung-Dao Lee Institute, Shanghai Jiao Tong University, Shanghai, 200240, China}
-\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, Toronto, ON M5S 3H4, Canada}
-\affiliation{Canadian Institute for Advanced Research, CIFAR Program in Gravitation and Cosmology, Toronto, ON M5G 1Z8, Canada}
-\affiliation{Perimeter Institute for Theoretical Physics, Waterloo, ON N2L 2Y5, Canada}
+\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, M5S 3H4, ON, Canada}
+\affiliation{Canadian Institute for Advanced Research, CIFAR Program in Gravitation and Cosmology, Toronto, M5G 1Z8, ON, Canada}
+\affiliation{Perimeter Institute for Theoretical Physics, Waterloo, N2L 2Y5, ON, Canada}
 
 \author{Xin~Wang}\email{\url{xwang@cita.utoronto.ca}}
-\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, M5S 3H8, Ontario, Canada}
+\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, M5S 3H8, ON, Canada}
 
@@
 These measurements are all parity-even.
-Here we show that, the presence of neutrino mass provides a unique contribution to the directions of the angular momentum  of galaxies, which is the first parity-odd neutrino effect on large scale structures.
-This parity-odd observable is free of the contamination of the linear perturbation theory, and can be cleanly separated from other non-gravitational effects. 
-A complete 21-cm neutral Hydrogen halo survey deep to redshift 1 can give a $5\sigma$ confidence level of detecting the neutrino torque effect if the sum of neutrino masses is 0.05 eV.
-
+Here we show that, the presence of neutrino mass provides a unique contribution to the directions of the angular momentum  of galaxies, which is the first parity-odd neutrino effect of galaxies or halos.
+This parity-odd observable is free of the contamination of linear perturbation theory, and can be cleanly separated from other non-gravitational effects.
+A complete 21-cm survey deep to redshift 1 can give a $5\sigma$ confidence level of detecting the neutrino torque effect if the sum of neutrino masses is 0.05 eV.
 \end{abstract}
@@
 
-\tcb{\textit{Introduction.---}}
+\textit{Introduction.---}
 Neutrino mass is a long-standing physics problem. 
@@
 $M_\nu \equiv \sum_{i=1}^3 m_{\nu_i} \gtrsim$ 0.05 eV \citep{2014ChPhC..38i0001O}. 
-The existence of neutrino mass has profound impacts on cosmic evolution. 
-In the early Universe relativistic neutrinos modulate the matter-to-radiation ratio, 
-and based on which the cosmic microwave background observations provide an 
+The existence of neutrino mass has profound impacts on cosmic evolution, and the current cosmic microwave background observations provide an 
 upper bound of $M_\nu\lesssim 0.23$ eV \citep{2016A&A...594A..13P}. 
-In the late Universe, neutrinos become non-relativistic, 
+In the cosmic structure formation, neutrinos are non-relativistic, 
 and contribute to the matter energy density $\Omega_m$. 
@@
 neutrinos maintain a high velocity dispersion, 
-known as ``free streaming'', which reduces their gravitational collapse on small scales. 
-A number of large scale structure (LSS) surveys 
-(such as Euclid \cite{2011arXiv1110.3193L},
-LSST \cite{2009arXiv0912.0201L}, eBOSS \cite{2016AJ....151...44D}, WFIRST\cite{2012SPIE.8442E..1UG}, 
-and DESI \cite{2015AAS...22533605E}) are aimed to improve this upper bound using neutrino effects on LSS, 
+known as ``free-streaming'', which reduces their gravitational collapse on small scales. 
+A number of large scale structure (LSS) surveys \citep{2011arXiv1110.3193L,2015AAS...22533605E}
+%(such as Euclid \cite{2011arXiv1110.3193L},
+%LSST \cite{2009arXiv0912.0201L}, eBOSS \cite{2016AJ....151...44D}, WFIRST\cite{2012SPIE.8442E..1UG}, 
+%and DESI \cite{2015AAS...22533605E})
+will improve this upper bound using neutrino effects on LSS, 
 however we usually encounter difficulties in disentangling neutrino effects from other parameters, 
@@
 For these effects one needs to carefully separate and exclude the contaminations from non-gravitational contributions. 
-The only clean, gravitational-only neutrino effect on LSS is the damping of weak gravitational lensing power spectrum \citep{2016A&A...594A..13P}. 
+The only clean, gravitational-only neutrino effect on LSS is the damping of weak gravitational lensing power spectrum \cite{2011arXiv1110.3193L}. 
 However the amount of information and constraining power of the lensing is limited by the projection onto the 2-dimensional sky. 
-Is there a 3-dimensional measure which is cleanly gravity-driven, to measure the neutrino mass?
+Is there a 3-dimensional measure that is cleanly gravity-driven, to measure the neutrino mass?
 
-The angular momentum. The angular momenta (spin) of dark matter halos or galaxies originated by clean gravitational interactions. 
-Unlike parity-even tidal and density fields, the linear perturbation theory (other effects and bias) cannot contaminate a parity-odd spin measurement. 
-Here we present the neutrino torque effect, to modulate the spin of dark matter halos.
+The angular momentum direction of galaxies. 
+The direction of angular momentum (hereafter spin) is readily observable, while the magnitude is not. 
+Majority of galaxies are disk galaxies, and the rotation axis is perpendicular to the disk \citep{1998MNRAS.297L..71M}. 
+The orientation of the disk is obtained by inclination and parallactic angles (parity-even) and dust absorption, and from the Doppler effect of spectral lines we determine the plus-minus sign of the spin, which is parity-odd.
+A parity-odd spin measurement cannot be contaminated by the linear perturbation theory. 
+Only gravity affects spin statistics of halos or galaxies at large separations, since baryons do not travel to the distance of the free-streaming of neutrinos.
 
-\tcb{\textit{Theory.---}} 
-In the picture of LSS formation, gravitational instability lets initial density fluctuations form dark matter halos, in which galaxies are embedded.
+We present the neutrino torque effect, to modulate the spin of dark matter halos.
+Here we use halos to represent galaxies.
+The galaxy spin can be observed to a much larger radii through the radio emission of the gas, and the spin change only very modestly to larger radii.
+This suggests that the spins of a galaxy and the whole halo are well correlated.
+
+\textit{Theory.---}
+In the picture of LSS formation, gravitational instability lets initial density fluctuations form dark matter halos, where galaxies are embedded.
 In these highly nonlinear structures, uncertainties of halo bias, 
@@
 
-In the tidal torque theory, the spin of a halo in Lagrangian space $\spin_L\propto\int_{V_L}\bs{q}\times\bs{\nabla}\phi\diff^3\bs{q}$ 
-can be written in first order as $\spin_I\propto\bs{\epsilon}\,\I\,\T$ \citep{1984ApJ...286...38W}, 
-where $\phi$ is the gravitational potential, $\bs{q}$ is the Lagrangian coordinates relative to the center of mass of the protohalo in volume $V_L$, 
-$\I_q=(I_{ij})=(\int_{V_L} q_i q_j \diff^3\bs{q})$ is the protogalactic inertia tensor, 
-$\T=(T_{ij})=(\partial_i\partial_j\phi)$ is the local tidal shear tensor, 
+The initial halo spin is written in Lagrangian space as 
+$\spin_L\propto-\int_{V_L}\bs{q}\times\bs{\nabla}\phi_c\diff^3\bs{q}$, 
+where $\phi_c$ is the gravitational potential of CDM, 
+$\bs{q}$ is the Lagrangian coordinates relative to the center of mass of the protohalo in volume $V_L$. 
+In the tidal torque theory \citep{1984ApJ...286...38W},
+it can be written as,
+$\spin_T\propto\bs{\epsilon}\,\I_q\,\T_c$, where
+$\I_q=(I_{ij})\equiv(\int_{V_L} q_i q_j \diff^3\bs{q})$,
+$\T_c=(T_{ij})\equiv(\partial_i\partial_j\phi_c)$
+are the protogalactic inertia tensor and the local tidal shear tensor\footnote{$\I_q$ 
+and $\T_c$ differ from textbook by trace, which does not contribute to $\spin_T$.}, 
 and $\bs{\epsilon}=(\epsilon_{ijk})$ is the Levi-Civita symbol to collect the asymmetric components generated by
-the misalignment between $\I_q$ and $\T$. This spin grows at first order in the early Universe \citep{1984ApJ...286...38W}. 
-Actually, as halos collapse, nonlinear gravity field is hard to torque small objects, and the spin directions are expected  to be contributed mostly in the linear regime.
+the misalignment between $\I_q$ and $\T_c$. Here, $\I_q$ and $\T_c$ are parity-even, $\bs{\epsilon}$ is parity-odd, and thus $\spin_T$ is parity-odd.
+In linear perturbation theory, $\phi_c$ remains constant, so $\T_c$ decays as $a^{-2}$ due to the cosmic expansion, where $a$ is the scale factor.
+%\footnote{$\phi_c(q,t)=\phi_c(q,t=0)$, $x=aq$, so $T_{x}=\partial^2_{x}\phi_c=\frac{1}{a^2}\partial^2_q\phi_c\propto\frac{1}{a^2}$.}
+Also, the protohalo shrinks in size and turns more spherical in shape.
+A decayed tidal field is hard to torque small and round objects, so the spin are expected to be contributed mostly in the linear regime.
 These concepts can be straightforwardly tested in numerical simulations. 
@@
 Massive neutrinos contribute sub-percent fraction of the matter ingredient, and their unique spatial distribution and evolution should contribute a unique torque to halos and galaxies.
-In particular, neutrino density field traces CDM on large scales while their small scale structures are smoothed out by their high velocity dispersion, known as the free-streaming. 
-They contribute a predictable tidal tensor $\T_\nu(m_{\nu_i})$ depending on the their mass. 
-The interplay between $\I_q$ and $\T_\nu$ leads to an additional neutrino torque $\spin^I_\nu\propto\bs{\epsilon}\,\I_q\,\T_\nu$.
-Integrated neutrino torque over cosmic evolution, we get the neutrino modulation of the halo spin,
-\begin{equation}\label{eq.jnu0}
-\spin^I_{\nu 0}=\int a(\tau)D(\tau)\bs{\epsilon}\I(\tau)\T_\nu(\tau)\diff\tau.
-\end{equation}
-Here, $\tau$ is the Newtonian time, $D(\tau)$ is the growth factor, and other quantities are functions of time.
+In particular, neutrino density field traces CDM on large scales while their small scale structures are smoothed out by their free-streaming. 
+They contribute a predictable tidal tensor $\T_\nu(m_{\nu_i})$ depending on the their mass \citep{2015PhRvD..92b3502I}. 
+The interplay between $\I_q$ and $\T_\nu$ leads to an additional neutrino torque $\spin^T_\nu\propto\bs{\epsilon}\,\I_q\,\T_\nu$.
+Integrated it over the cosmic evolution, we denote the total neutrino modulation of the halo spin as $\spin^T_{\nu 0}$ given by the tidal torque theory.
+%\begin{equation}\label{eq.jnu0}
+%\spin^I_{\nu 0}=\int a(\tau)D(\tau)\bs{\epsilon}\I(\tau)\T_\nu(\tau)\diff\tau.
+%\end{equation}
+%Here, $\tau$ is the Newtonian time, $D(\tau)$ is the growth factor, and other quantities are functions of time.
 
-\tcb{\textit{Reconstruction.---}} 
-The feasibility of prediction of the neutrino tidal torque relies on the precision reconstruction of the cosmic evolution history. 
-There are many emerging reconstruction methods. 
-For example, the isobaric halo reconstruction can recover the initial conditions of the universe from spatial distribution of halos. 
-In the case of high number densities the reconstructed density field is correlated with the true initial density field on scales $k\lesssim 0.7 h/{\rm Mpc}$ \citep{2017ApJ...847..110Y}, 
-close to the limit ($k\lesssim 1 h/{\rm Mpc}$) of isobaric reconstruction from using direct CDM density field \citep{2017PhRvD..96l3502Z}, 
+\textit{Reconstruction.---}
+The feasibility of predicting neutrino torque relies on the precision reconstruction of $\I_q$ and $\T_\nu$.
+
+Neutrino distribution shares the same Fourier phases of CDM from halo reconstruction \citep{2015PhRvD..92b3502I,2017ApJ...847..110Y}, but only differs by the ratio of their linear transfer function, depending on the neutrino masses.
+The reconstruction $\T_\nu$ is reliable even if it is applied directly on halos \citep{2015PhRvD..92b3502I}, because they are relatively linear even at present epoch.
+
+The reconstruction of $\I_q$ relies on the fact that $\I_q$ and $\T_c$ are highly correlated in Lagrangian space \citep{2000ApJ...532L...5L,2001ApJ...555..106L},
+which can be understood that $\I_q$ is a collection of matter to be shell-crossed to form a halo, being parallel with $\T_c$.
+The latter depends on a precise reconstruction of CDM initial conditions. 
+Recent years, many emerging reconstruction methods have achieved unprecedented accuracies.
+For example, the isobaric halo reconstruction \citep{2017PhRvD..96l3502Z} can recover the initial conditions of the Universe from spatial distribution of halos.
+In the case of high halo number densities the reconstructed density field is correlated with the true initial density field on scales $k\lesssim 0.7 h/{\rm Mpc}$ \citep{2017ApJ...847..110Y}, 
+close to the limit ($k\lesssim 1 h/{\rm Mpc}$) of isobaric reconstruction from using direct CDM density field \citep{2017MNRAS.469.1968P}, 
 or the limit of reconstruction from the true displacement field \citep{2017PhRvD..95d3501Y}. 
-ELUCID simulation are able to reconstruct the local Universe \citep{2014ApJ...794...94W}.
+ELUCID simulations are able to reconstruct the full evolution history of the real local Universe \citep{2014ApJ...794...94W}.
 These reconstruction techniques enable us to study the tidal field from both CDM and neutrinos in an unprecedented precision, at different epochs of the cosmic evolution.
-Neutrinos share the same Fourier phases of CDM, reconstructed from halo reconstruction, but only differ by the ratio of their linear transfer function, depending on the neutrino masses.
-The final neutrino torque can be understood as the interaction between large (neutrino free streaming) and small (collapsing) scales.
+We construct an equivalent inertia $\I_R$ from $\T_c$, in order to maximize the cross-correlation $\mu$ between $\bs{\epsilon}\,\I_R\T_\nu$ and $\spin^T_{\nu 0}$ (see Appendix \ref{app.I}).
 
-To apply Eq.(\ref{eq.jnu0}), $\I_q$ is also required to be reconstructed. 
-Firstly, we note the fact that, during the initial conditions, the $\I_q$ and $\T_c$ are highly correlated in Lagrangian space \citep{2000ApJ...532L...5L,2001ApJ...555..106L}, 
-which can be understood that $\I_q$ is a collection of matter to be shell-crossed to form a halo, being parallel with $\T_c$.
-Secondly, in the linear, intermediate epochs, $\I_q$ is firstly reshaped according to the dominating $\T_c$, thus the linear evolved $\T_\nu(\tau)$ will act on an evolved $\I(\tau)$.
-Further, $\I(\tau)$ will be affected by nonlinear effects, which are more difficult to predict, however halos are relatively small and their spin directions can hardly be changed dramatically.
-\tcb{A deeply study of the second step above is beyond the scope of this letter. Here we show that, even simplified to the first point above, the total neutrino torque can still be reconstructed.}
-
-We construct an equivalent inertia $\I_r$ from $\T_c$, in order that $\bs{\epsilon}\,\I_r\T_\nu$ is maximumly correlated with $\spin^I_{\nu 0}$ from Eq.(\ref{eq.jnu0}).
-In the primary coordinate of $\T_c$, $\T_c$ can be eigen-decomposed as $\T_c=\sum_{i=1}^3\T^{\lambda_i}_c$ and find $\alpha_i$ such that the reconstructed
-\begin{equation}\label{eq.Ir}
-	\I_r=\sum_{i=1}^3\alpha_i\T^{\lambda_i}_c
-\end{equation}
-optimizes the cross-correlation $\mu(\spin^I_{\nu 0},\spin^{I_r})$, where
-\begin{equation}\label{eq.jnuIr}
- 	\spin^{I_r}_\nu=\bs{\epsilon}\,\I_r\T_\nu
-\end{equation}
-is the reconstructed neutrino spin direction.
+%$\T_\nu$ can be considered as a low-pass window function convolution with galaxy distribution and $\T_c$ is used as a proxy of $\I_q$. 
+The final neutrino torque can be understood as an interaction between two scales -- a small collapsing scale $\T_c$,
+a large neutrino free-streaming scale $\T_\nu$,
+and with an antisymmetric (parity-odd) operator $\bs{\epsilon}$ collecting the antisymmetric (parity-odd) contributions from the multiplication of these two symmetric (parity-even) tensors.
 
@@
  while dotted ellipsoids show the tidal shear from CDM and neutrinos $\T_c$ and $\T_\nu$. $\spin_L$, 
- $\spin_0$ are the Lagrangian and Eulerian halo spins, whereas $\spin_I$ is the tidal torque prediction. 
- The initial neutrino torque is shown by $\spin_\nu$.}\label{fig.1}
+ $\spin_0$ are the Lagrangian and Eulerian halo spins, whereas $\spin_T$ is the tidal torque prediction. 
+ The initial neutrino torque is shown by $\spin^T_\nu$.}\label{fig.1}
 \end{figure}
-\tcb{\textit{Simulation.---}}
+
+\textit{Simulation.---}
 These correlations and coefficients are tested across a set of high-resolution $N$-body simulations \citep{2018ApJS..237...24Y}. 
@@
 we select a very massive halo ($7.8\times 10^{14}M_\odot$) to maximize the clarity of the visualization of the halo properties.
-Note that these properties (cross-correlations) have only weak dependence on halo mass. 
+We confirm in simulation that these properties (cross-correlations) have only weak dependence on halo mass. 
 The background LSS at redshift $z=0$ has the thickness $2r_{200}$ with $r_{200}$ being the halo radius within which the mean halo density is 200 times the mean matter density of the Universe.
 The Lagrangian mapping of this halo is shown by the protohalo's column density with the orange clouds.
-To visualize the tidal torque theory, we plot ellipsoids equivalent to $\I_q$ and 
+To visualize the tidal torque theory, we plot equivalent ellipsoids (solid curves) with their moment of inertia equal to $\I_q$ and 
 $\I_{200}$, where $\I_{200}$ is the moment of inertia within $r_{200}$. 
 The ellipsoids with dotted lines correspond to $\T_c$ and $\T_\nu$, normalized such that their volumes are $V_L$ and $8V_L$ respectively. 
-As expected, $\I_q$ and $\T_c$ are aligned with their primary axes in parallel with the collapsing direction, or perpendicular to the filament below. 
-Their minor misalignment yields the tidal torque $\spin_I$ (all the spin arrows are normalized to have 15 Mpc$/h$), which is the first order approximation of the true initial spin $\spin_L$. 
+As expected, $\I_q$ and $\T_c$ are aligned with their primary axes in parallel with the collapsing direction,  perpendicular to the filament containing the halo. 
+Their minor misalignment yields the tidal torque $\spin_T$, which is the first order approximation of the true initial spin $\spin_L$ (all the spin arrows are normalized to have 15 Mpc$/h$). 
 They are, in general, highly correlated with the spin of the final halo $\spin_0$.
-In comparison, the neutrino tidal shear $\T_\nu$ torques $\I_q$ in an other less correlated direction $\spin_\nu$.
+In comparison, the neutrino tidal shear $\T_\nu$ torques $\I_q$ in an other less correlated direction $\spin^T_\nu$.
 
 The validity of the tidal torque formulation is tested by an ensemble average over all halos, 
-across 3 orders of magnitude in mass range, and over simulations with different resolutions.
-the cross-correlation coefficients $\left\langle \spin(z) \cdot \spin_L \right\rangle$ and
-$\left\langle \spin(z) \cdot \spin_I \right\rangle$ smoothly decrease from 1 to 0.80, and
+across 3 orders of magnitude in mass range ($>2\times10^{12}M_\odot$), and over simulations with different resolutions.
+As the Universe evolves from the initial condition to $z=0$, the cross-correlation coefficients $\left\langle \spin(z) \cdot \spin_L \right\rangle$ and
+$\left\langle \spin(z) \cdot \spin_T \right\rangle$ smoothly decrease from 1 to 0.80, and
 from 0.75 to 0.69, respectively.
 
-For neutrinos, the first-order tidal torque approximation gives a perfect (with cross-correlation 0.99) representation of the actual neutrino torque and it has generally less than 0.2 cross-correlation with CDM torques.
-This is expected in that $\T_c$ dominates locally \tcr{(power spectrum index?)} whereas $\T_\nu$ is contributed beyond the neutrino free-streaming scale \tcr{(is this a 4-point function?)}.
+For neutrinos, the first-order tidal torque approximation gives a near perfect (with cross-correlation 0.99) representation of the actual neutrino torque, and it has generally $<0.2$ cross-correlates with CDM torques.
+This is expected in that $\T_c$ dominates locally  whereas $\T_\nu$ is contributed beyond the neutrino free-streaming scale.
 These two species, however, have a highly correlated contribution in structure formation.
-When we consider the forces that the two species exerted to the protohalo, $\bs{F}_{c/\nu}\propto\int_{V_L}\bs{\nabla}\phi_{c/\nu}\diff^3\bs{q}$, the cross-correlation between two species is as high as 0.86 \tcr{(power spectrum index?)}.
+When we consider the gravitational forces that the two species exerted to the protohalo, $\bs{F}_{c/\nu}\propto\int_{V_L}\bs{\nabla}\phi_{c/\nu}\diff^3\bs{q}$, the cross-correlation between two species is as high as 0.86 .
 
-By Eq.(\ref{eq.jnu0}) we estimate the magnitude of integrated neutrino torque $\left\langle|\spin_{0\nu}|/|\spin_0|\right\rangle\simeq 3\times10^{-4}$.
-In particular, the effect given by the smoother distribution for neutrinos relative to CDM accounts 0.03, while the neutrino fraction $f_\nu=3.5\times 10^{-3}$ (for $M_\nu=0.05$ eV) and the backreaction from neutrinos to CDM \tcr{$\sqrt{8}$} contribute the rest.
+From simulations, we estimate the magnitude of integrated neutrino torque $\left\langle|\spin^T_{\nu 0}|/|\spin_0|\right\rangle\simeq 3\times10^{-4}$.
+In particular, the effect given by the smoother distribution for neutrinos relative to CDM accounts 0.03, while the neutrino fraction $f_\nu=3.5\times 10^{-3}$ (for $M_\nu=0.05$ eV) and the back-reaction factor from neutrinos to CDM $\sqrt{8}$ \citep{2012MNRAS.420.2551B} contribute the rest.
 
-Measured from simulations, $(\alpha_1,\alpha_2,\alpha_3)=(-0.18,0.18,0.02)$ in Eq.(\ref{eq.Ir}) optimizes the cross-correlation $\left\langle\spin^I_{\nu 0}\cdot\spin^{I_r}_\nu\right\rangle=0.19$.
+Measured from simulations, the cross-correlation between $\bs{\epsilon}\,\I_R\T_\nu$ and $\spin^T_{\nu 0}$ is about $\mu=0.19$ (see Appendix \ref{app.I}).
 We then need $5\times 10^9$ halos to have a $5\sigma$ detection (see Appendix \ref{app.error}).
-If we improve the reconstruction of $\I_R$, the lower limit of required halos is $2\times 10^8$.
+If we improve the reconstruction of $\I_R$ from studying the cosmic evolution history \citep{2014ApJ...794...94W}, the lower limit of required halos is $2\times 10^8$.
 
-\tcb{\textit{Discussion.---}} 
-Accuracy and mass dependency: we notice that the cross-correlation coefficients between $\spin_0$ and initial values $\spin_L$ and $\spin_T$ are prominently higher than what was measured in \citep{2000ApJ...532L...5L}.
-We investigate that numerical errors that may affect the results.
+\textit{Discussion.---}
+We found that the cross-correlation coefficients between $\spin_0$ and initial values $\spin_L$ and $\spin_T$ are prominently higher than our previous understandings (e.g.
+\citep{2000ApJ...532L...5L}).
+We carefully investigate the numerical errors that may affect the results. 
 P3M (particle-particle particle-mesh) algorithms result in higher cross-correlation between initial and final spins, compared to PM (particle-mesh), where additional tangential forces in PM violate the angular momentum conservation.
-Higher mass halos in a given simulation generally have slightly higher $\mu_L$ and $\mu_I$, however the correlation is enhanced greatly as we use higher mass resolutions.
+Higher mass halos in a given simulation generally have slightly higher cross-correlations between initial and final spins, however the correlation is enhanced as we use higher mass resolutions. 
 All other cross-correlation measurements have only weak dependencies on the halo mass, even in a fixed simulation. 
-With different box sizes, mass resolutions, force resolutions, and find that the results are consistent across these simulations.
-Especially, in the estimation of number of halo spins needed to detect the neutrino torque, the deciding number is the cross-correlation $\left\langle\spin^I_{\nu 0}\cdot\spin^{I_r}_\nu\right\rangle=0.19$.
+With different box sizes, mass resolutions, force resolutions, we find that the results are consistent across these simulations. 
+The estimation of number of halo spins needed to detect the neutrino torque depends on $\mu=0.19$.
 This has a very weak dependence on halo mass and configuration of the simulation.
-An accurate study of the neutrino torque requires mass and force resolutions to cover the wide range of halo mass, a large box size ($>600\,{\rm Mpc}/h$) to account the neutrino tides at distance, and neutrino particles or fluids to calculate more precise nonlinear neutrino effects.
-These require a simulation comparable to the computing resources of TianNu \citep{2017RAA....17...85E}.
+An accurate study of the neutrino torque requires mass and force resolutions to cover the wide range of halo mass, a large box size ($>600\,{\rm Mpc}/h$) to account the neutrino tides at distance, and neutrino particles/fluids, further studies of reconstruction of $\I_R$ (Appendix \ref{app.I}) to calculate more precise nonlinear neutrino effects on an evolving halo.
+These require future simulations with computing power comparable to that of TianNu \citep{2017RAA....17...85E}.
 
-Observability: surveys like the Hubble Sphere Hydrogen Survey (HSHS) \citep{2006astro.ph..6104P} can cover 1 billion galaxies. There are xxx HI galaxies with redshift 1 \citep{2004MNRAS.350.1210Z}.
+Surveys like the Hubble Sphere Hydrogen Survey (HSHS) \citep{2006astro.ph..6104P} or a modified 21cm Cosmic Vision \citep{2018arXiv180207216D} are able to observe order of $10^9$ HI galaxies ($>10^{12}M_\odot$),
+in a cosmic volume $(4{\rm Gpc}/h)^3$ below redshift $z\simeq 1$ \citep{2004MNRAS.350.1210Z}. 
+Under the standard model of the Universe with $M_\nu=0.05$ eV, a $5\sigma$ detection will be reachable.
+Beyond standard models, e.g., the neutrino mass could be generated by a gravitational $\theta$-term condensation as the source of dark energy \citep{2016PhRvD..93k3002D}, the tidal torque history will be different, and we need different numbers of galaxies to differentiate between these models.
 
-Beyond standard models: it was also suggested that neutrino mass is generated by a gravitational $\theta$-term condensation as the source of dark energy \citep{2016PhRvD..93k3002D}.
+%Codis \citep{2018MNRAS.tmp.2447C}.
 
-Codis \citep{2018MNRAS.tmp.2447C}.
-
-\tcb{\textit{Conclusion.---}} 
+\textit{Conclusion.---}
 Enormous efforts has been contributed to the neutrino mass properties.
@@
 The angular momentum is a 3-dimensional, gravity-driven, parity-odd measure, 
-and is easily observable, well modeled by the tidal torque theory, and well conserved over the cosmic evolution.
+and is readily observable, well modeled by the tidal torque theory, and well conserved over the cosmic evolution.
 It contains comparable amount of information as density field but poorly appreciated in application.
-The recently developed reconstruction techniques enable us to visit an-order-of-magnitude more precise initial conditions of the Universe in Lagrangian space.
+Recent developed reconstruction techniques enable us to visit an-order-of-magnitude more precise initial conditions of the Universe in Lagrangian space, 
+to accurately reconstruct the neutrino torque effects,
+and to unveil the neutrino mass properties in upcoming galaxy surveys.
 
-\tcb{\textit{Acknowledgments.---}}
+\textit{Acknowledgments.---}
 We acknowledge funding from NSERC.
+The simulations were performed on the Sunnyvale cluster at CITA and on the Niagara supercomputer at the SciNet HPC Consortium.
 
@@
 \appendix
-\section{Errors in 3D}\label{app.error}
-There are $N$ halos with their unit spin vector randomly distributed on a 2D sphere, $|\spin|=1$ and $\left\langle \spin \right\rangle =\bs{0}$.
-Adding an additional vector $a\hat{\bs{x}}$ ($a\ll 1$) to $\spin$ and normalize we have $\spin'=(\spin+a\hat{\bs{x}})/|\spin+a\hat{\bs{x}}|$, then we project $\spin'$ onto $\hat{\bs{x}}$ and we get the signal $b=\spin'\cdot\hat{\bs{x}}$.
-The numerical statistics are as following: $\left\langle b \right\rangle=2a/3$ and $\sigma(b)=1/\sqrt{3N}$.
-So, to get a $n\sigma$ detection,
-\begin{equation}
-	N=3n^2/4a^2 
-\end{equation}
-halos will be needed.
+\section{Reconstruction of $\I_R$}\label{app.I}
+In Lagrangian space, $\I_q$ and $\T_c$ are highly correlated.
+In linear, intermediate epochs, $\I_q$ is firstly reshaped according to the dominating $\T_c$, thus the linear evolved $\T_\nu(\tau)$ will act on an evolved $\I(\tau)$.
+Further, $\I(\tau)$ will be affected by nonlinear effects, which are more difficult to predict, however halos are relatively small and their spin directions can hardly be changed dramatically.
+A deep study of the latter steps above is beyond the scope of this letter. Here we show that, even simplified to the first point above, the total neutrino torque can still be reconstructed.
+
+We construct an equivalent inertia $\I_R$ from $\T_c$, in order to maximum the cross-correlation coefficient $\mu$ between $\bs{\epsilon}\,\I_R\T_\nu$ and $\spin^I_{\nu 0}$.
+In the primary coordinate of $\T_c$, $\T_c$ can be eigen-decomposed as $\T_c=\sum_{i=1}^3\T^{\lambda_i}_c$ and we find $(\alpha_1,\alpha_2,\alpha_3)=(-0.7,0.7,0.08)$ (normalized such that $\sum \alpha_i^2=1$) and $\I_R\propto\sum_{i=1}^3\alpha_i\T^{\lambda_i}_c$
+optimize $\mu$ to be 0.19.
+
+\section{Errors}\label{app.error}
+Consider $N$ halos with their unit spin vector randomly distributed on a 2-dimensional sphere, i.e., $|\spin|=1$ and $\left\langle \spin \right\rangle =\bs{0}$.
+Adding an additional vector $\epsilon\hat{\bs{x}}$ ($\epsilon\ll 1$) to $\spin$ and normalize, $\spin'=(\spin+\epsilon\hat{\bs{x}})/|\spin+\epsilon\hat{\bs{x}}|$, then project $\spin'$ onto $\hat{\bs{x}}$ and we get $p=\spin'\cdot\hat{\bs{x}}$.
+Since $\left\langle p \right\rangle=2\epsilon/3$ and $\sigma(p)=1/\sqrt{3N}$, an $n\sigma$ detection requires $N=3n^2/4\epsilon^2$.
 

### 2018-12-29 reply_to_editor+referee_v2.1.txt
modification v1.0

Reply to the editor

Dear Editor,

Thank you for your consideration of our manuscript and thank the referee to give valuable comments. We think we can can address all the questions. Especially for the major comment, the referee should look at section 4.2 of 1804.07306 and their conclusion that spin 'direction' correlation is strong. Fig.2 of 1804.07306 shows the uncorrelated spin 'amplitude' correlation, which is unrelated to our work. We also explained it to the referee below.

We think our manuscript appropriate for PRL is that we proposed a new approach and observable to measure the neutrino mass, which has not been considered before. Most proposed neutrino observables are not proven to work, including weak lensing. It is too early to prove that this method works, but rather it is meant to encourage further studies in this new field of angular momentum directions.

Sincerely,

Authors

Reply to the referee

Dear referee,

Thank you for your comments and suggestions in our paper. We answer your comments below and marked the modifications in our paper.

We are proposing a totally new approach to measure neutrino mass, and we have proven that there is a new observable that has not been considered before. Most proposed neutrino observables are not proven to work, not even CMB lensing because of the baryon feedback. The importance of our paper is that it is completely new and is meant to encourage new research studies on the angular momentum connection.

----------------------------------------------------------------------
Report of Referee A -- LX16362/Yu
----------------------------------------------------------------------

Major comments:

My biggest concern with this work is that I think the authors need to
assume that the angular momentum of galaxies traces the angular
momentum of halos. I have done a quick search in the literature and
found 1804.07306, where in Fig. 2 I can see a large scatter between
the spin of galaxies and the spin of their host halos. The authors
need to address this point in detail and explain how they will
distinguish the expected small neutrino angular momentum contribution
from the underlying scatter.

>> Answer:
We agree that the spin connection between galaxies and halos is an important concern. Most literature study the alignment of shapes, or only the amplitude of the angular momentum. Most parts of 1804.07306 focus on the ‘amplitude’ of the angular momentum.
In 1804.07306, is says “It is intriguing that the amplitudes of spins are uncorrelated although the alignment of the spin vectors is relatively good”. In Fig. 2 of 1804.07306, it shows the correlation between angular momentum ‘amplitudes’, not the correlation between ‘directions’. In 1804.07306, Section 4.2, 5.2 and Figure 4, 7 discussed the spin ‘direction’ correlations, and their most important conclusion is that “the median cos\theta_{gas,halo} is 0.71 to 0.73”. This suggests a lower limit of the spin direction correlation between DM halos and baryons, since in higher resolution simulations the angular momentum direction is better conserved. Although 1804.07306 did not study the Eulerian-Lagrangian spin direction correlation, from results from our work it indicates that stars and gas correlates should also correlate with Lagrangian spin direction. The importance of our paper means to start new research studies on the angular momentum connection.
We modified the last part of the introduction clearer — “Recent hydrodynamic galaxy formation simulations suggest that, although uncorrelated in amplitude, the spin directions of stellar, cold gas components and CDM halo are strongly correlated (e.g. section 4.2 of \citep{2018arXiv180407306J}).”

Intermediate comments:

The authors argue that the spin of a galaxy changes very modestly with
distance, so they relate halo spin to galaxy spin. However, this can
not be true in galaxies clusters, where different galaxies inside the
cluster will have different angular momentum. The authors should
explain how this measurement will be done in practice. They should
also explain how they will deal with satellites galaxies.

>> Answer:
Thanks for this comment. 1804.07306 uses only zoom-in simulations and lower mass halos. In ref \citep{2016MNRAS.460.3772S} they use cosmological simulations and found that the central galaxy, DM halo and the distribution of satellite galaxies are aligned with their minor axis of their moments of inertia. The minor axis of the moment of inertia is an indicator of spin axis, however this requires further studies. Related to our work, in clusters most galaxies are elliptical which we are not considering. Only 1% of galaxies are in clusters, so it is not a major effect.
In the first paragraph of discussion, we added “We do not propose to use galaxies in clusters, and most spiral galaxies are not in clusters.”
Our paper is not trying to prove the applicability, but to propose that there is a new neutrino observable.  And we have proven that there is a new observable that has not been considered before. Most proposed neutrino observables are not ‘proven’ to work, not even CMB lensing (also because of baryon feedback).  The importance of the paper is that it is completely new, and is meant to start new research studies on the angular momentum connection.

The authors state that “A decayed tidal field is hard to torque small
and round objects, so the spin are expected to be contributed mostly
in the linear regime”. While I agree with this statement, I think the
authors can not neglect the contribution of non-linear and baryonic
effects, that although are expected to be small, so is the signal from
neutrinos. The authors write that CDM and baryons are torqued by the
same gravitational shear, such as the spin direction is not affected
by baryonic effects. I agree with the authors for halos, but not for
individual galaxies, where AGN feedback, dynamic friction…etc can
potentially affect the galaxy properties.

>> Answer:
We agree and try to describe in the text more clearly.
Firstly, “A decayed tidal field is hard to torque small and round objects, so the spin are expected to be contributed mostly in the linear regime” is to physically and qualitatively explain why the halo spin between Lagrangian space, Eulerian space and tidal torque theory is as high as 0.8, and we rewrote the CDM-baryon relations in the next paragraph separately, as —
“In Lagrangian space, CDM and baryonic matter are torqued by the same gravitational shear, so their subsequent local evolution should not systematically change their respective spin directions. For example, baryonic feedback has to conserve angular momentum, unless the baryons are expelled from the galaxy and carry angular momentum away systematically. The dynamical friction between baryons and dark matter can only increase the angular momentum correlation. These all indicate a high correlation between CDM and baryons (e.g. \citep{2018arXiv180407306J}).”

I agree with the authors that neutrinos should contribute to the total
angular momentum of a halo by a predictable amount. However, I think
the Lagrangian region enclosing the neutrinos that end up in the halo
will be very large, due to the neutrinos thermal velocities. How do
the authors deal with that?

>> Answer:
Yes. We agree. Since initially only neutrinos have large velocity dispersion, not CDM and baryons, so the Lagrangian region for neutrinos (enclosed in the final halo region) is very large, while CDM and baryons should be equivalent. Note that the neutrinos enclosed in the halo region are not necessarily gravitationally bound and they do not evolve with CDM and baryons, so their spin (indeed very small) should not contribute to the CDM halo and baryons. Even considering it, the neutrino faction is small, and the spin is small, so it is a second order effect. The effect neutrino tides torques the CDM+baryon inertia is what we consider.
We rewrote the sentence as — “Massive neutrinos contribute sub-percent fraction of the matter ingredient, and their unique spatial distribution and evolution should contribute a unique torque to CDM and baryons.”

Small comments:

Introduction: update neutrino constraints from Planck 2018

>> Answer: Thanks. We updated it.

When the authors mention: “In the cosmic structure formation,
neutrinos are non-relativistic”, they should state that this is a low
redshift, not at high-redshift, were neutrinos are relativistic.

>> Answer: Thanks. We modify the sentence as — “At low redshifts, neutrinos become non-relativistic, and contribute to…”

I do not agree with the authors that a clean observable is the effect
of neutrinos on the weak lensing power spectrum. There are a large set
of systematics that affect these observations, and from the theory
side it is still unclear the contamination by baryonic effects.

>> Answer: 
Weak lensing effect is purely gravitational. We agree with you that it is not clean in the systematics and baryonic effects. We modify the sentence as — “Thus most popular cosmological neutrino mass measurements are not proven to work. For example, although the neutrino effect on weak lensing power spectrum \citep{2011arXiv1110.3193L} uses pure gravity, there are a large set of systematics affecting the observations and unclear baryonic effects that can contaminate the neutrino signal.”

>> Thank you again for all the comments and suggestions. All the modified parts in the paper are marked with blue text.

Authors

### 2018-12-29 reply_to_referee_v2.0.txt
modification v1.0

--- 2018-12-29 reply_to_editor+referee_v2.1.txt
@@
-Reply to the editor
-
-Dear Editor,
-
-Thank you for your consideration of our manuscript and thank the referee to give valuable comments.
-We think we can can address all the questions.
-Especially for the major comment, the referee should look at section 4.2 of 1804.07306 and their conclusion that spin 'direction' correlation is strong.
-Fig.2 of 1804.07306 shows the uncorrelated spin 'amplitude' correlation, which is unrelated to our work.
-We also explained it to the referee below.
-
-We think our manuscript appropriate for PRL is that we proposed a new approach and observable to measure the neutrino mass, which has not been considered before.
-Most proposed neutrino observables are not proven to work, including weak lensing.
-It is too early to prove that this method works, but rather it is meant to encourage further studies in this new field of angular momentum directions.
-
-Sincerely,
-
-Authors
-
 Reply to the referee
-
-Dear referee,
-
-Thank you for your comments and suggestions in our paper. We answer your comments below and marked the modifications in our paper.
-
-We are proposing a totally new approach to measure neutrino mass, and we have proven that there is a new observable that has not been considered before.
-Most proposed neutrino observables are not proven to work, not even CMB lensing because of the baryon feedback.
-The importance of our paper is that it is completely new and is meant to encourage new research studies on the angular momentum connection.
 
@@
 Only 1% of galaxies are in clusters, so it is not a major effect.
-In the first paragraph of discussion, we added “We do not propose to use galaxies in clusters, and most spiral galaxies are not in clusters.”
+In the last part of the introduction we added “ In galaxy clusters, central galaxies and satellites are also found to be aligned with halos \citep{2016MNRAS.460.3772S}.”
+And, in the first paragraph of discussion, we added “For galaxy clusters, the correlation between minor axes in \cite{2016MNRAS.460.3772S} also implies the correlation between spins.
+After all, however, since most galaxies are not in clusters so the contribution of clusters is not a major effect.”
 Our paper is not trying to prove the applicability, but to propose that there is a new neutrino observable.
@@
 
->> Answer: 
-Weak lensing effect is purely gravitational.
-We agree with you that it is not clean in the systematics and baryonic effects.
-We modify the sentence as — “Thus most popular cosmological neutrino mass measurements are not proven to work.
-For example, although the neutrino effect on weak lensing power spectrum \citep{2011arXiv1110.3193L} uses pure gravity, there are a large set of systematics affecting the observations and unclear baryonic effects that can contaminate the neutrino signal.”
+>> Answer: We agree with it.
+We modify the sentence as — “For example, although the neutrino effect on weak lensing power spectrum \citep{2011arXiv1110.3193L} uses pure gravity, there are a large set of systematics affecting the observations and unclear baryonic effects that can contaminate the neutrino signal.”
 
 >> Thank you again for all the comments and suggestions. All the modified parts in the paper are marked with blue text.
-
 Authors

### 2018-12-29 neu_torque.tex
typos
modification v1.0

@@
 \begin{abstract}
-Cosmological observations are promising ways to improve of our understandings of the neutrino mass properties.
+Cosmological observations are promising ways to improve our understandings of the neutrino mass properties.
 The upper bound on their sum of mass is given by the cosmic microwave background and large scale structures.
@@
 % insert suggested PACS numbers in braces on next line
-\pacs{}
+\pacs{98.80.-k}
 % insert suggested keywords - APS authors don't need to do this
@@
 The existence of neutrino mass has profound impacts on cosmic evolution, and the current cosmic microwave background observations provide an 
-upper bound of $M_\nu\lesssim 0.23$ eV \citep{2016A&A...594A..13P}. 
-In the cosmic structure formation, neutrinos are non-relativistic, 
-and contribute to the matter energy density $\Omega_m$. 
+upper bound of \tcb{ $M_\nu\lesssim 0.12$ eV \citep{2018arXiv180706209P}}. 
+\tcb{At low redshifts, neutrinos become non-relativistic,} 
+and contribute to the matter energy density $\Omega_m$ in the structure formation. 
 Unlike the majority of matter, the cold dark matter (CDM) and baryons, 
 neutrinos maintain a high velocity dispersion, 
-known as ``free-streaming'', which reduces their gravitational collapse on small scales. 
+referred to as ``free-streaming'', which reduces their gravitational collapse on small scales. 
 A number of large scale structure (LSS) surveys \citep{2011arXiv1110.3193L,2015AAS...22533605E}
@@
 %and DESI \cite{2015AAS...22533605E})
-will improve this upper bound using neutrino effects on LSS, 
-however we usually encounter difficulties in disentangling neutrino effects from other parameters, 
+plan to improve this upper bound using neutrino effects on LSS,
+however we usually encounter difficulties in disentangling it from other non-gravitational effects, 
 and in understanding halo bias and cosmic variance. 
-Recently, new nonlinear neutrinos effects on LSS are proposed to give independent measurements on neutrino mass \citep{2014PhRvL.113m1301Z,2016PhRvL.116n1301Z,2017NatAs...1E.143Y}. 
-For these effects one needs to carefully separate and exclude the contaminations from non-gravitational contributions. 
-The only clean, gravitational-only neutrino effect on LSS is the damping of weak gravitational lensing power spectrum \cite{2011arXiv1110.3193L}. 
-However the amount of information and constraining power of the lensing is limited by the projection onto the 2-dimensional sky. 
+\tcb{Thus most popular cosmological neutrino mass measurements are not proven to work.
+For example, although the neutrino effect on weak lensing power spectrum \citep{2011arXiv1110.3193L} uses pure gravity, 
+there are a large set of systematics affecting the observations and unclear baryonic effects that can contaminate the neutrino signal.}
+Besides, the amount of information and constraining power of the lensing is limited by the projection onto the 2-dimensional sky. 
 Is there a 3-dimensional measure that is cleanly gravity-driven, to measure the neutrino mass?
+%Many new neutrino effects on LSS are proposed \citep{2014PhRvL.113m1301Z, 2016PhRvL.116n1301Z, 2017NatAs...1E.143Y}.
 
 The angular momentum direction of galaxies. 
-The direction of angular momentum (hereafter spin) is readily observable, while the magnitude is not. 
+The \tcb{\it direction} of angular momentum (hereafter \tcb{\it spin}) is readily observable, while the magnitude is not. 
 Majority of galaxies are disk galaxies, and the rotation axis is perpendicular to the disk \citep{1998MNRAS.297L..71M}. 
@@
 Here we use halos to represent galaxies.
-The galaxy spin can be observed to a much larger radii through the radio emission of the gas, and the spin change only very modestly to larger radii.
-This suggests that the spins of a galaxy and the whole halo are well correlated.
+The galaxy spin can be observed to a much larger radii through the radio emission of the gas, and the spin change only very modestly to larger radii \citep{wiki:Galaxy_rotation_curve}. 
+\tcb{Recent hydrodynamic galaxy formation simulations suggest that, although uncorrelated in amplitude, the spin directions of stellar, 
+cold gas components and CDM halo are strongly correlated (e.g. section 4.2 of \citep{2018arXiv180407306J}). 
+%In galaxy clusters, central galaxies and satellites are also found to be aligned with halos \citep{2016MNRAS.460.3772S}.
+}
 
@@
 Also, the protohalo shrinks in size and turns more spherical in shape.
-A decayed tidal field is hard to torque small and round objects, so the spin are expected to be contributed mostly in the linear regime.
-These concepts can be straightforwardly tested in numerical simulations. 
-In Lagrangian space, CDM and baryonic matter are torqued by the same gravitational shear, so the spin direction is not affected by baryonic effects. 
+A decayed tidal field is hard to torque small and round objects, so the spin are expected to be contributed mostly in the linear regime \citep{2002MNRAS.332..325P}.
 If a halo has a merging history, it simply collects disconnected regions in Lagrangian space but conserves their total spin.
+These concepts can be straightforwardly tested in $N$-body simulations.
 
-Massive neutrinos contribute sub-percent fraction of the matter ingredient, and their unique spatial distribution and evolution should contribute a unique torque to halos and galaxies.
+In Lagrangian space, CDM and baryonic matter are torqued by the same gravitational shear, so 
+\tcb{their subsequent local evolution should not systematically change their respective spin directions.
+For example, baryonic feedback has to conserve angular momentum, unless the baryons are expelled from the galaxy and carry angular momentum away systematically.
+The dynamical friction between baryons and dark matter can only increase the angular momentum correlation.
+These all indicate a high correlation between CDM and baryons (e.g. \citep{2018arXiv180407306J}).}
+
+Massive neutrinos contribute sub-percent fraction of the matter ingredient, and their unique spatial distribution and evolution should contribute a unique torque to \tcb{CDM and baryons}.
 In particular, neutrino density field traces CDM on large scales while their small scale structures are smoothed out by their free-streaming. 
@@
 %Here, $\tau$ is the Newtonian time, $D(\tau)$ is the growth factor, and other quantities are functions of time.
+\tcb{
+This torque is unique in that, the neutrino free-streaming scale is much larger than the scale to which local nonlinear and baryonic effects can systematically contribute.
+}
 
@@
 \textit{Discussion.---}
+\tcb{
+We proposed a new approach to measure the neutrino mass that has not been considered before. 
+The applicability relies on the new research studies on the angular momentum connection.
+Galaxy formation simulations like \cite{2018arXiv180407306J,2016MNRAS.460.3772S} are needed to study the Lagrangian spin correlations for baryons, 
+despite that combining \cite{2018arXiv180407306J} and our result already implies a strong correlation. 
+%For galaxy clusters, the correlation between minor axes in \cite{2016MNRAS.460.3772S} also implies the correlation between spins.
+After all, however, since most galaxies are not in clusters so the contribution of clusters is not a major effect.
+We do not propose to use galaxies in clusters, and most spiral galaxies are not in clusters.
+}
+
 We found that the cross-correlation coefficients between $\spin_0$ and initial values $\spin_L$ and $\spin_T$ are prominently higher than our previous understandings (e.g.
@@
 An accurate study of the neutrino torque requires mass and force resolutions to cover the wide range of halo mass, a large box size ($>600\,{\rm Mpc}/h$) to account the neutrino tides at distance, and neutrino particles/fluids, further studies of reconstruction of $\I_R$ (Appendix \ref{app.I}) to calculate more precise nonlinear neutrino effects on an evolving halo.
-These require future simulations with computing power comparable to that of TianNu \citep{2017RAA....17...85E}.
+These require future simulations with computing power comparable to that of TianNu \citep{2017NatAs...1E.143Y,2017RAA....17...85E}.
 
@@
 Under the standard model of the Universe with $M_\nu=0.05$ eV, a $5\sigma$ detection will be reachable.
-Beyond standard models, e.g., the neutrino mass could be generated by a gravitational $\theta$-term condensation as the source of dark energy \citep{2016PhRvD..93k3002D}, the tidal torque history will be different, and we need different numbers of galaxies to differentiate between these models.
+Beyond standard models, e.g., the neutrino mass could be
+generated through a neutrino vacuum condensate triggered by a gravitational $\theta$-term
+\citep{2016PhRvD..93k3002D}, the tidal torque history will be different, 
+and we need different numbers of galaxies to differentiate between these models.
 
@@
 Enormous efforts has been contributed to the neutrino mass properties.
-Radio astronomic and cosmological surveys are promising low-cost experiments implemented in the Universe to probe basic physic mysteries. 
+Radio astronomic and cosmological surveys are promising low-cost experiments implemented in the Universe to probe basic physics mysteries. 
 The angular momentum is a 3-dimensional, gravity-driven, parity-odd measure, 
@@
 \textit{Acknowledgments.---}
+\tcb{We thank the anonymous referee who gives valuable comments which improved this letter.} 
+HRY thanks Derek Inman, Yipeng Jing, Pengjie Zhang and Jiaxin Han for helpful comments and suggestions.
 We acknowledge funding from NSERC.

### 2019-03-26 neu_torque.tex
2nd modification from ref.

@@
 \begin{abstract}
-Cosmological observations are promising ways to improve our understandings of the neutrino mass properties.
-The upper bound on their sum of mass is given by the cosmic microwave background and large scale structures.
-These measurements are all parity-even.
-Here we show that, the presence of neutrino mass provides a unique contribution to the directions of the angular momentum  of galaxies, which is the first parity-odd neutrino effect of galaxies or halos.
-This parity-odd observable is free of the contamination of linear perturbation theory, and can be cleanly separated from other non-gravitational effects.
-A complete 21-cm survey deep to redshift 1 can give a $5\sigma$ confidence level of detecting the neutrino torque effect if the sum of neutrino masses is 0.05 eV.
+Cosmological observations are promising ways to improve our understanding of neutrino mass properties.
+The upper bound on the sum of masses is given by the cosmic microwave background and large scale structure.
+These measurements are all parity-even, and potentially contaminated by unmodelled baryonic effects.
+In this paper we propose a novel partity-odd gravitational effect of neutrinos: A unique contribution to the directions of the angular momentum field of galaxies and halos.
+This observable is free of contamination in linear perturbation theory, and thus likely more cleanly separated from other non-gravitational effects.
+A deep 21-cm survey to redshift 1 can potentially yield a $5\sigma$ significance on neutrino mass detection for a fiducial sum of neutrino masses of 0.05 eV.
 \end{abstract}
@@
 
-\textit{Introduction.---}
+\section{Introduction} 
 Neutrino mass is a long-standing physics problem. 
@@
 $M_\nu \equiv \sum_{i=1}^3 m_{\nu_i} \gtrsim$ 0.05 eV \citep{2014ChPhC..38i0001O}. 
-The existence of neutrino mass has profound impacts on cosmic evolution, and the current cosmic microwave background observations provide an 
-upper bound of \tcb{ $M_\nu\lesssim 0.12$ eV \citep{2018arXiv180706209P}}. 
-\tcb{At low redshifts, neutrinos become non-relativistic,} 
+The existence of neutrino mass has profound impacts on cosmic evolution, and the current cosmic microwave background (CMB) observations provide an 
+upper bound of $M_\nu\lesssim 0.12$ eV \citep{2018arXiv180706209P}. 
+At low redshifts, neutrinos become non-relativistic, 
 and contribute to the matter energy density $\Omega_m$ in the structure formation. 
@@
 %and DESI \cite{2015AAS...22533605E})
-plan to improve this upper bound using neutrino effects on LSS,
-however we usually encounter difficulties in disentangling it from other non-gravitational effects, 
-and in understanding halo bias and cosmic variance. 
-\tcb{Thus most popular cosmological neutrino mass measurements are not proven to work.
-For example, although the neutrino effect on weak lensing power spectrum \citep{2011arXiv1110.3193L} uses pure gravity, 
-there are a large set of systematics affecting the observations and unclear baryonic effects that can contaminate the neutrino signal.}
-Besides, the amount of information and constraining power of the lensing is limited by the projection onto the 2-dimensional sky. 
-Is there a 3-dimensional measure that is cleanly gravity-driven, to measure the neutrino mass?
+plan to improve this upper bound using neutrino effects on LSS.
+Their approaches either measure the impact of neutrinos on the growth of structure by measurement at different redshifts, or on different length scales.
+
+The small gravitational contribution from neutrinos requires a precise measurement at two different epochs or scales.
+Most techniques aim to combine the CMB with a second, lower redshift, and use small scales where plenty of modes are available for a precise measurement.
+To measure the impact on scale dependence requires the measurement on a larger scale, where cosmic variance limits the number of observable modes.
+Here we propose to exploit a historic {\it fossil} comparison at large scales: we measure the {\it same} modes at two redshifts in the full 3-D LSS volume, which overcomes the cosmic variance limit.
+Neutrinos free stream, and are mostly smooth below their free streaming scale, tens of Mpc.
+Baryons are conserved, and even in the most extreme feedback scenarios do not move more than 10 Mpc relative to CDM.
+In this paper we present a novel approach to exploit the torquing effect by neutrinos to probe their mass.
+This is meant as a first exploration, intended to stimulate observational and more detailed theoretical studies.
+%In section \ref{sec.spin} ...
+
+%however we usually encounter difficulties in disentangling it from other non-gravitational effects, 
+%and in understanding halo bias and cosmic variance. 
+%\tcb{Thus most popular cosmological neutrino mass measurements are not proven to work.
+%For example, although the neutrino effect on weak lensing power spectrum \citep{2011arXiv1110.3193L} uses pure gravity, 
+%there are a large set of systematics affecting the observations and unclear baryonic effects that can contaminate the neutrino signal.}
+%Besides, the amount of information and constraining power of the lensing is limited by the projection onto the 2-dimensional sky. 
+%Is there a 3-dimensional measure that is cleanly gravity-driven, to measure the neutrino mass?
 %Many new neutrino effects on LSS are proposed \citep{2014PhRvL.113m1301Z, 2016PhRvL.116n1301Z, 2017NatAs...1E.143Y}.
 
-The angular momentum direction of galaxies. 
-The \tcb{\it direction} of angular momentum (hereafter \tcb{\it spin}) is readily observable, while the magnitude is not. 
+\section{Spin of galaxies}\label{sec.spin}
+The {\it direction} of angular momentum (hereafter {\it spin}) of a galaxy is readily observable, while the magnitude is not. 
 Majority of galaxies are disk galaxies, and the rotation axis is perpendicular to the disk \citep{1998MNRAS.297L..71M}. 
@@
 Here we use halos to represent galaxies.
-The galaxy spin can be observed to a much larger radii through the radio emission of the gas, and the spin change only very modestly to larger radii \citep{wiki:Galaxy_rotation_curve}. 
-\tcb{Recent hydrodynamic galaxy formation simulations suggest that, although uncorrelated in amplitude, the spin directions of stellar, 
-cold gas components and CDM halo are strongly correlated (e.g. section 4.2 of \citep{2018arXiv180407306J}). 
+The galaxy spin can be observed to a much larger radii through the 21-cm radio emission of cold gas, and the spin change only very modestly to larger radii \citep{2001ARA&A..39..137S}. 
+Hydrodynamic galaxy formation simulations suggest that spin directions of stellar, 
+cold gas components, and CDM halo are correlated (e.g. \citep{2010MNRAS.404.1137B,2010MNRAS.405..274H,2011MNRAS.415.2607D,2015ApJ...812...29T} and references therein)\footnote{We
+note that although the angular momentum \emph{magnitudes} of dark matter halos and
+baryons are not correlated, their angular momentum \emph{directions} (spins) are 
+highly correlated (e.g. section 4.2 of \citep{2018arXiv180407306J}), with a median
+misalignment angle typically $30^\circ$.}.
+The magnitudes are not as correlated \citep{2018arXiv180407306J}, as one might expect from the
+substantial dissipation and outflow.
+The observed mostly co-planar rotation of baryons in late type galaxies indicates that additional dissipation or selective outflow are unlikely to change the angular momentum direction,
+which is common to most baryons, while the magnitude is likely to change, and varies within the baryonic inventory.
 %In galaxy clusters, central galaxies and satellites are also found to be aligned with halos \citep{2016MNRAS.460.3772S}.
-}
 
-\textit{Theory.---}
+\section{Theory}\label{sec.theory}
 In the picture of LSS formation, gravitational instability lets initial density fluctuations form dark matter halos, where galaxies are embedded.
@@
 In Lagrangian space, CDM and baryonic matter are torqued by the same gravitational shear, so 
-\tcb{their subsequent local evolution should not systematically change their respective spin directions.
-For example, baryonic feedback has to conserve angular momentum, unless the baryons are expelled from the galaxy and carry angular momentum away systematically.
+their subsequent local evolution should not systematically change their respective spin directions.
+For example, baryonic feedback has to conserve angular momentum, unless the baryons are expelled from the galaxy with systematic misaligned angular momentum.
 The dynamical friction between baryons and dark matter can only increase the angular momentum correlation.
-These all indicate a high correlation between CDM and baryons (e.g. \citep{2018arXiv180407306J}).}
+These all indicate a high correlation between CDM and baryons (e.g. \citep{2010MNRAS.404.1137B,2010MNRAS.405..274H,2011MNRAS.415.2607D,2015ApJ...812...29T,2018arXiv180407306J}).
 
-Massive neutrinos contribute sub-percent fraction of the matter ingredient, and their unique spatial distribution and evolution should contribute a unique torque to \tcb{CDM and baryons}.
+Massive neutrinos contribute sub-percent fraction of the matter ingredient, and their unique spatial distribution and evolution should contribute a unique torque to CDM and baryons.
 In particular, neutrino density field traces CDM on large scales while their small scale structures are smoothed out by their free-streaming. 
@@
 %Here, $\tau$ is the Newtonian time, $D(\tau)$ is the growth factor, and other quantities are functions of time.
-\tcb{
 This torque is unique in that, the neutrino free-streaming scale is much larger than the scale to which local nonlinear and baryonic effects can systematically contribute.
-}
 
-\textit{Reconstruction.---}
+\section{Reconstruction}\label{sec.reco}
 The feasibility of predicting neutrino torque relies on the precision reconstruction of $\I_q$ and $\T_\nu$.
@@
 Recent years, many emerging reconstruction methods have achieved unprecedented accuracies.
-For example, the isobaric halo reconstruction \citep{2017PhRvD..96l3502Z} can recover the initial conditions of the Universe from spatial distribution of halos.
+For example, the isobaric halo reconstruction \citep{2017PhRvD..96l3502Z,2018arXiv180706381W} can recover the initial conditions of the Universe from spatial distribution of halos.
 In the case of high halo number densities the reconstructed density field is correlated with the true initial density field on scales $k\lesssim 0.7 h/{\rm Mpc}$ \citep{2017ApJ...847..110Y}, 
@@
 
-\textit{Simulation.---}
+\section{Simulation}\label{label.sim}
 These correlations and coefficients are tested across a set of high-resolution $N$-body simulations \citep{2018ApJS..237...24Y}. 
@@
 
-In Fig.\ref{fig.1} (all quantities are projected onto the plane of this letter), 
+In Fig.\ref{fig.1} (all quantities are projected onto the plane of this paper), 
 we select a very massive halo ($7.8\times 10^{14}M_\odot$) to maximize the clarity of the visualization of the halo properties.
@@
 
-\textit{Discussion.---}
-\tcb{
-We proposed a new approach to measure the neutrino mass that has not been considered before. 
+\section{Discussion}\label{sec.discussion}
+We proposed a novel approach to measure the neutrino mass that has not been considered before. 
+Two independent observables are compared to infer neutrino mass:
+the displacement field measures the inertia tensor on small scales, 
+and the neutrino tidal field on large scales.
+This is compared to the angular momentum field on the same large scales, 
+thus not subject to cosmic variance.
+The two fields measure the tides at two different epochs: spin is a fossil of the epoch of galaxy formation.
 The applicability relies on the new research studies on the angular momentum connection.
-Galaxy formation simulations like \cite{2018arXiv180407306J,2016MNRAS.460.3772S} are needed to study the Lagrangian spin correlations for baryons, 
-despite that combining \cite{2018arXiv180407306J} and our result already implies a strong correlation. 
+Galaxy formation simulations are needed to study the Lagrangian spin correlations for baryons, 
+despite that combining \cite{2010MNRAS.404.1137B,2010MNRAS.405..274H,2011MNRAS.415.2607D,2015ApJ...812...29T,2018arXiv180407306J} and \citep{2002MNRAS.332..325P} already implies a strong correlation. 
+We do not propose to use galaxies in clusters, and most spiral galaxies are not in clusters.
 %For galaxy clusters, the correlation between minor axes in \cite{2016MNRAS.460.3772S} also implies the correlation between spins.
 After all, however, since most galaxies are not in clusters so the contribution of clusters is not a major effect.
-We do not propose to use galaxies in clusters, and most spiral galaxies are not in clusters.
-}
 
@@
 An accurate study of the neutrino torque requires mass and force resolutions to cover the wide range of halo mass, a large box size ($>600\,{\rm Mpc}/h$) to account the neutrino tides at distance, and neutrino particles/fluids, further studies of reconstruction of $\I_R$ (Appendix \ref{app.I}) to calculate more precise nonlinear neutrino effects on an evolving halo.
-These require future simulations with computing power comparable to that of TianNu \citep{2017NatAs...1E.143Y,2017RAA....17...85E}.
+These require future simulations with computing power comparable to that of TianNu \citep{2017RAA....17...85E}.
 
@@
 
-\textit{Conclusion.---}
-Enormous efforts has been contributed to the neutrino mass properties.
+\section{Conclusion}\label{sec.conclusion}
+Enormous efforts has been contributed to the neutrino mass properties and new, independent cosmological neutrino mass effects are proposed to constrain the neutrino mass \citep{2014PhRvL.113m1301Z,2016PhRvL.116n1301Z,2017NatAs...1E.143Y,2017PhRvD..95h3518I,2019PhRvL.122d1302C}.
 Radio astronomic and cosmological surveys are promising low-cost experiments implemented in the Universe to probe basic physics mysteries. 
@@
 
-\textit{Acknowledgments.---}
-\tcb{We thank the anonymous referee who gives valuable comments which improved this letter.} 
+\acknowledgments
+We thank the anonymous referee who gives valuable comments which improved this paper.
 HRY thanks Derek Inman, Yipeng Jing, Pengjie Zhang and Jiaxin Han for helpful comments and suggestions.
@@
 Further, $\I(\tau)$ will be affected by nonlinear effects, which are more difficult to predict, however halos are relatively small and their spin directions can hardly be changed dramatically.
-A deep study of the latter steps above is beyond the scope of this letter. Here we show that, even simplified to the first point above, the total neutrino torque can still be reconstructed.
+A deep study of the latter steps above is beyond the scope of this paper. Here we show that, even simplified to the first point above, the total neutrino torque can still be reconstructed.
 

### 2019-03-30 neu_torque.tex
more refs

@@
 The applicability relies on the new research studies on the angular momentum connection.
-Galaxy formation simulations are needed to study the Lagrangian spin correlations for baryons, 
-despite that combining \cite{2010MNRAS.404.1137B,2010MNRAS.405..274H,2011MNRAS.415.2607D,2015ApJ...812...29T,2018arXiv180407306J} and \citep{2002MNRAS.332..325P} already implies a strong correlation. 
+
+From past and current $N$-body simulations, the spin correlation between halos and their Lagrangian protohalos  
+(e.g. \citep{2000ApJ...532L...5L,2002MNRAS.332..325P} and this work) is higher than 0.5, 
+meaning that the nonlinear effects at low redshifts only add less than 
+factor of 2 scatter to the spin correlation.
+Recent galaxies formation simulations
+(e.g. \cite{2010MNRAS.404.1137B,2010MNRAS.405..274H,2011MNRAS.415.2607D,2015ApJ...812...29T,2018arXiv180407306J}) also show a $\sim 0.5$ spin correlation between galaxies and halos,
+and thus baryonic effects also contribute only factor of 2.
+These strong correlations are much higher than the $\sim 0.01$ order correlation between galaxy shapes
+and true tidal fields by weak lensing observations \citep{2013JCAP...01..026A}. 
 We do not propose to use galaxies in clusters, and most spiral galaxies are not in clusters.
